---
title: "Transcriptomics Analysis Example"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Setup

## Load Libraries

```{r load-libraries}
library(tidyverse)
library(readxl)
library(DESeq2)
library(sva)
library(pheatmap)
library(ggplot2)

# Source custom functions
source("transcriptomics_functions.R")
```

# Method 1: Using the Complete Pipeline

This is the simplest approach - one function call does everything:

```{r pipeline-simple}
# Run complete pipeline
results <- run_deseq2_pipeline(
  counts_path = './output/count_data_all.csv',
  metadata_path = "data/metadata_ecphage.xlsx",
  annotation_path = "./output/22ET500456__with-ecolik12.csv",
  design_formula = "~ time_point + antibiotic + phage + ab_conc",
  min_count = 10,
  use_sva = TRUE,
  metadata_filter_col = "comment",
  metadata_filter_pattern = "missing",
  fname_col = "fname"
)

# Extract components
counts <- results$counts
metadata <- results$metadata
annotation <- results$annotation
dds <- results$dds
vsd <- results$vsd
svobj <- results$sva
```

# Method 2: Step-by-Step Approach

If you want more control, use individual functions:

```{r step-by-step, eval=FALSE}
# Step 1: Load data
data_list <- load_transcriptomics_data(
  counts_path = './output/count_data_all.csv',
  metadata_path = "data/metadata_ecphage.xlsx",
  annotation_path = "./output/22ET500456__with-ecolik12.csv",
  metadata_filter_col = "comment",
  metadata_filter_pattern = "missing",
  fname_col = "fname"
)

counts <- data_list$counts
metadata <- data_list$metadata
annotation <- data_list$annotation

# Step 2: Create DESeq2 object with SVA
deseq_results <- create_deseq2_dataset(
  counts = counts,
  metadata = metadata,
  design_formula = "~ time_point + antibiotic + phage + ab_conc",
  min_count = 10,
  use_sva = TRUE,
  sva_formula = "~ time_point + phage + ab_conc",  # Custom SVA formula
  blind = TRUE
)

dds <- deseq_results$dds
vsd <- deseq_results$vsd
svobj <- deseq_results$sva
```

# Data Summary

```{r data-summary}
# Quick summary
summarize_deseq2(dds)

# Metadata overview
cat("\n=== Metadata Variables ===\n")
str(metadata)

# Sample distribution
cat("\n=== Sample Distribution ===\n")
table(metadata$time_point, metadata$phage)
```

# Quality Control

## PCA Plot

```{r pca-plot, fig.width=10, fig.height=6}
# PCA with different color schemes
pcaData <- plotPCA(vsd, intgroup = c("time_point", "phage"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

ggplot(pcaData, aes(x = PC1, y = PC2, color = time_point, shape = phage)) +
  geom_point(size = 3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  theme_bw() +
  theme(legend.position = "right") +
  ggtitle("PCA - VST Transformed Data")
```

## Sample Distances

```{r sample-distances, fig.width=10, fig.height=8}
# Calculate sample distances
sampleDists <- dist(t(assay(vsd)))
sampleDistMatrix <- as.matrix(sampleDists)

# Create annotation for heatmap
annotation_col <- data.frame(
  TimePoint = metadata$time_point,
  Phage = metadata$phage,
  Antibiotic = metadata$antibiotic,
  row.names = colnames(vsd)
)

# Plot heatmap
pheatmap(
  sampleDistMatrix,
  clustering_distance_rows = sampleDists,
  clustering_distance_cols = sampleDists,
  annotation_col = annotation_col,
  main = "Sample-to-Sample Distances"
)
```

## Dispersion Plot

```{r dispersion-plot, fig.width=8, fig.height=6}
plotDispEsts(dds, main = "Dispersion Estimates")
```

# Differential Expression Analysis

## Define Contrasts

```{r define-contrasts}
# Example: Compare phage treatment vs control at specific time point
# Adjust according to your experimental design

# View all coefficients
resultsNames(dds)
```

## Extract Results

```{r extract-results}
# Example contrast - modify based on your design
res <- results(dds, 
               contrast = c("phage", "treated", "control"),  # Adjust levels
               alpha = 0.05)

# Summary
summary(res)

# Order by adjusted p-value
res_ordered <- res[order(res$padj), ]

# View top genes
head(res_ordered, 10)
```

## Volcano Plot

```{r volcano-plot, fig.width=8, fig.height=6}
# Prepare data for plotting
res_df <- as.data.frame(res) %>%
  mutate(
    gene = rownames(res),
    significant = padj < 0.05 & abs(log2FoldChange) > 1,
    direction = case_when(
      padj < 0.05 & log2FoldChange > 1 ~ "Up",
      padj < 0.05 & log2FoldChange < -1 ~ "Down",
      TRUE ~ "NS"
    )
  )

# Create volcano plot
ggplot(res_df, aes(x = log2FoldChange, y = -log10(padj), color = direction)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = c("Up" = "red", "Down" = "blue", "NS" = "grey")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", alpha = 0.5) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Volcano Plot",
    x = "Log2 Fold Change",
    y = "-Log10 Adjusted P-value",
    color = "Regulation"
  )
```

## MA Plot

```{r ma-plot, fig.width=8, fig.height=6}
plotMA(res, main = "MA Plot", ylim = c(-5, 5))
```

## Top Differentially Expressed Genes

```{r top-genes}
# Get significant genes
sig_genes <- subset(res_ordered, padj < 0.05 & abs(log2FoldChange) > 1)

cat("Number of significant DEGs:", nrow(sig_genes), "\n")
cat("  - Upregulated:", sum(sig_genes$log2FoldChange > 0), "\n")
cat("  - Downregulated:", sum(sig_genes$log2FoldChange < 0), "\n")

# Show top 20
head(as.data.frame(sig_genes), 20)
```

## Heatmap of Top Genes

```{r top-genes-heatmap, fig.width=10, fig.height=8}
# Select top 50 genes by adjusted p-value
top_genes <- head(rownames(res_ordered), 50)

# Get normalized counts
norm_counts <- get_normalized_counts(dds, method = "vst", vsd = vsd)

# Subset and scale
mat_scaled <- t(scale(t(norm_counts[top_genes, ])))

# Plot
pheatmap(
  mat_scaled,
  annotation_col = annotation_col,
  show_rownames = TRUE,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  main = "Top 50 Differentially Expressed Genes (Z-score)"
)
```

# Export Results

```{r export-results}
# Create results table with annotations
if (!is.null(annotation)) {
  # Merge with annotation if available
  res_export <- as.data.frame(res_ordered) %>%
    rownames_to_column("gene_id") %>%
    left_join(annotation, by = c("gene_id" = "Geneid"))  # Adjust join column as needed
} else {
  res_export <- as.data.frame(res_ordered) %>%
    rownames_to_column("gene_id")
}

# Save results
write.csv(res_export, 
          file = "DESeq2_results.csv", 
          row.names = FALSE)

# Save significant genes only
sig_res <- res_export %>%
  filter(padj < 0.05 & abs(log2FoldChange) > 1)

write.csv(sig_res,
          file = "DESeq2_significant_genes.csv",
          row.names = FALSE)

cat("Results exported to CSV files\n")
```

# Session Info

```{r session-info}
sessionInfo()
```
