---
title: "IMC Heat Flow Curve Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 150)
```

```{r libraries}
rm(list=ls())
library(tidyverse)
library(patchwork)
library(knitr)
library(DescTools)

theme_tk <- theme_bw(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "#F0F0F0"),
    plot.title       = element_text(face = "bold", size = 13)
  )

# ============================================================
# USER PARAMETERS
# ============================================================

signals_file  <- "data/calscreener/expriments_PAS_signals_long.csv"
metadata_file <- "data/calscreener/expriments_PAS_well_metadata.csv"
specs_file    <- "data/calscreener/device_specs.csv"

# Curves before this time are dominated by the instrument's insertion/
# equilibration transient (large positive/negative swings unrelated to
# bacterial metabolism) and are excluded from plots and derived metrics.
startup_exclude_h <- 1

# Bin width (h) used to downsample the ~2s-resolution raw signal for
# plotting and summary statistics.
bin_width_h <- 0.1

# Colors per condition. AB 8 mg/L is kept (it is the isolate's known
# broth-microdilution ATM-AVI MIC, so the IMC-based MIC confirmation section
# needs it). AB 16 mg/L is excluded (see Data Import note): redundant above
# the known MIC, and was the source of the original interval-ratio-method
# instability that led to replacing that method.
curve_colors <- c(
  "Bacteria"            = "#2C3E50",   # Dark grey
  "AB 8 mg/L"           = "#922B21",   # Dark red
  "AB 2 mg/L"           = "#E67E22",   # Orange
  "AB 1 mg/L"           = "#F39C12",   # Amber
  "AB 0.5 mg/L"         = "#FAD7A0",   # Light amber
  "Phage"               = "#2980B9",   # Blue
  "Media ctrl"          = "#BDC3C7",   # Grey
  "Phage + AB 2 mg/L"   = "#8E44AD",   # Purple  <- highest-dose combo available
  "Phage + AB 1 mg/L"   = "#BB8FCE",   # Medium purple
  "Phage + AB 0.5 mg/L" = "#D7BDE2"    # Light purple
)

# Line types
curve_lty <- c(
  "Bacteria"            = "solid",
  "AB 8 mg/L"           = "dashed",
  "AB 2 mg/L"           = "dashed",
  "AB 1 mg/L"           = "dashed",
  "AB 0.5 mg/L"         = "dashed",
  "Phage"               = "dotdash",
  "Media ctrl"          = "dotted",
  "Phage + AB 2 mg/L"   = "solid",     # Highlight - highest-dose combo available
  "Phage + AB 1 mg/L"   = "solid",
  "Phage + AB 0.5 mg/L" = "solid"
)
```

## Data Import & Cleaning

> Data are the real CalScreener export (`data/calscreener/`), replacing the
> earlier workflow that read heat-flow curves digitized by hand from
> screenshots of the CalScreener software
> (`data/calscreener/old/heat_flow_digitized_0.25h_from_heat_png.csv`, no
> longer used). This batch is a single 48-well experiment
> ("Hanna 456 10P1 2 1 0.5 repeat 10-17-2023") with triplicate wells per
> condition (duplicate for `CTL MH`), and reference wells (rows A and F)
> excluded from analysis.
>
> **Note:** this batch's phage+antibiotic combination wells only go up to
> 2 mg/L ATM-AVI (0.5 / 1 / 2 mg/L). It does not include a 4 mg/L or 8 mg/L
> phage+antibiotic combination well, so this dataset cannot support any
> claim about phage + 4 mg/L specifically (e.g. matching the manuscript's
> 4 mg/L transcriptomics condition) -- that requires a separate IMC run.
> Statements below are restricted to the available 0.5-2 mg/L combination
> data.
>
> **AB 16 mg/L is excluded from this analysis entirely.** It is redundant
> above the isolate's known broth-microdilution ATM-AVI MIC (8 mg/L) and was
> also the concentration where a first attempt at IMC-based MIC confirmation
> (a 30-minute-interval heat ratio, following Tellapragada et al.) was least
> stable (see "Antibiotic-Only MIC Confirmation" below for what replaced that
> method). **AB 8 mg/L is retained** because it is the isolate's known MIC
> and is needed there to check the IMC-derived result against it.
>
> **Processing order** (raw -> calibration check -> baseline correction ->
> smoothing -> replicate aggregation -> optional relative normalization),
> reconstructing what calView itself computed rather than assuming a
> transformation:
>
> 1. `signal_raw` (not the earlier `signal_uW` convenience column) is the
>    starting measurement for every well.
> 2. **Calibration check**: `device_specs.csv` provides a per-channel
>    `gain`/`offset_W`. As an external validation (documented in this
>    project's calibration audit, not re-run inside this Rmd), five candidate
>    transformations of `signal_raw` (raw as-is; +offset; x gain;
>    x gain + offset; (+offset) x gain) were each averaged over that well's
>    own stored baseline window (`baseline_start_s`-`baseline_end_s`) and
>    compared against calView's own saved `baseline_average_signal` across
>    all 48 wells, and independently cross-checked against calView's saved
>    `main_integral_value`. **Result: `signal_raw` used as-is reproduces both
>    quantities almost exactly** (mean |reconstructed - xml_baseline| ~2e-8 W
>    across 48 wells; integral ratio within 5% of 1.0 for 37/41 wells with
>    non-negligible baselines -- 4 exceptions noted below), while every
>    transform involving `gain`/`offset_W` was off by 2-3 orders of
>    magnitude. This means `signal_raw` is already calView's calibrated
>    signal (gain/offset already applied upstream) -- applying
>    `device_specs.csv` again here would double-calibrate, so only the
>    single check below (reproducing `baseline_average_signal`) is re-run in
>    this Rmd; `specs` (loaded from `device_specs.csv`) is otherwise unused.
>    **4 wells did not reproduce `main_integral_value` within 5%** (C6: ratio
>    ~1.08; C7: ~1.96; D1: ~0.70; D3: ~0.74) -- two are Phage + 2 mg/L
>    replicate wells (C6/C7) and two are AB 1 mg/L replicate wells (D1/D3).
>    The cause was not resolved (most likely a small difference between this
>    Rmd's trapezoidal integration and calView's internal integration at the
>    window edges); results for these four wells' conditions should be read
>    with that residual uncertainty in mind.
> 3. **Baseline correction** (a different thing from calibration): each
>    well's own `baseline_average_signal` is subtracted from its raw signal,
>    matching how calView's `main_integral_value` is itself defined.
> 4. Reference wells (rows A/F, `is_reference = TRUE`) are excluded from
>    biological analysis (kept only for QC) -- per-well baseline correction
>    is used instead of subtracting a reference well, since the device
>    operates in differential mode and reference handling is assumed to
>    already be reflected in the exported signal.
> 5. Only after per-well calibration and baseline correction are replicate
>    wells averaged (never average raw signal across replicates first).
>
> **Replicates are technical, not biological:** all wells in this batch come
> from one CalScreener experiment (one plate, one run); the n=2-3 "replicate
> wells" per condition are technical replicate wells within that single
> experiment, not independent biological experiments. This is stated
> explicitly wherever replicate counts are reported below, and is why no
> inferential statistics (p-values) are computed on them (see Descriptive
> Comparisons).

```{r data_prep}
meta <- read_csv(metadata_file, show_col_types = FALSE)
signals <- read_csv(signals_file, show_col_types = FALSE)
specs <- read_csv(specs_file, show_col_types = FALSE)

# ------------------------------------------------------------------
# Step 1-2: calibration check. Confirms signal_raw needs no gain/offset
# transform before it is used as the heat-flow measurement (see note above).
# ------------------------------------------------------------------
baseline_windows <- meta %>%
  select(well, baseline_start_s, baseline_end_s, baseline_average_signal)

calib_check <- signals %>%
  select(well, time_s, signal_raw) %>%
  inner_join(baseline_windows, by = "well") %>%
  filter(time_s >= baseline_start_s, time_s <= baseline_end_s) %>%
  summarise(recon_raw = mean(signal_raw),
            baseline_average_signal = first(baseline_average_signal),
            .by = well) %>%
  mutate(abs_diff = abs(recon_raw - baseline_average_signal))

cat("Calibration check (signal_raw vs. calView baseline_average_signal):\n")
cat("  Mean |diff| across", nrow(calib_check), "wells:", mean(calib_check$abs_diff), "\n")
cat("  Max  |diff|:", max(calib_check$abs_diff), "\n")
cat("  -> signal_raw used as-is; no gain/offset applied.\n")

# ------------------------------------------------------------------
# Second validation: reproduce calView's own main-section integral
# (IntegralValue) as the trapezoidal integral of the baseline-corrected
# signal over the stored main window. Confirms the baseline-correction
# step (not just the raw signal itself) matches calView's definition.
# ------------------------------------------------------------------
main_windows <- meta %>%
  select(well, main_start_s, main_end_s, main_integral_value, baseline_average_signal)

integral_check <- signals %>%
  select(well, time_s, signal_raw) %>%
  inner_join(main_windows, by = "well") %>%
  filter(time_s >= main_start_s, time_s <= main_end_s) %>%
  arrange(well, time_s) %>%
  summarise(
    auc_baseline_sub = DescTools::AUC(time_s, signal_raw - first(baseline_average_signal),
                                       method = "trapezoid"),
    main_integral_value = first(main_integral_value),
    .by = well
  ) %>%
  filter(abs(main_integral_value) > 0.05) %>%  # exclude near-zero denominators
  mutate(ratio = main_integral_value / auc_baseline_sub)

cat("\nIntegral check (baseline-corrected AUC vs. calView main_integral_value):\n")
cat("  Wells tested (|integral| > 0.05):", nrow(integral_check), "\n")
cat("  Median ratio (xml / reconstructed):", median(integral_check$ratio, na.rm = TRUE), "\n")
cat("  Wells within 5% of ratio 1.0:",
    sum(abs(integral_check$ratio - 1) < 0.05, na.rm = TRUE), "/", nrow(integral_check), "\n")
cat("  -> baseline-corrected signal_raw, integrated over calView's stored\n")
cat("     main window, reproduces calView's own IntegralValue.\n")
cat("  Wells outside 5% (unresolved discrepancy):\n")
integral_check %>%
  filter(abs(ratio - 1) >= 0.05) %>%
  select(well, main_integral_value, auc_baseline_sub, ratio) %>%
  print()

# Well -> condition mapping, derived from the real sample names recorded by
# the CalScreener software (see metadata `sample_name` / `repeat_group`).
# CTL BACT 16 (AB 16 mg/L) is intentionally left out of this lookup -- see
# Data Import note -- so its wells are dropped by the inner_join below
# (sample_name has no match -> row is dropped).
condition_lookup <- tribble(
  ~sample_name,     ~condition,              ~group,
  "CTL BACT",       "Bacteria",              "Bacteria",
  "CTL BACT 0.5",   "AB 0.5 mg/L",           "AB only",
  "CTL BACT 1",     "AB 1 mg/L",             "AB only",
  "CTL BACT 2",     "AB 2 mg/L",             "AB only",
  "CTL BACT 8",     "AB 8 mg/L",             "AB only",
  "CTL EC10P1",     "Phage",                 "Phage alone",
  "CTL MH",         "Media ctrl",            "Control",
  "10P1 0.5MG",     "Phage + AB 0.5 mg/L",   "Phage + AB",
  "10P1 1MG",       "Phage + AB 1 mg/L",     "Phage + AB",
  "10P1 2MG",       "Phage + AB 2 mg/L",     "Phage + AB"
)

meta_use <- meta %>%
  filter(!is_reference) %>%
  select(well, sample_name, repeat_group, baseline_average_signal) %>%
  inner_join(condition_lookup, by = "sample_name") %>%
  mutate(
    ab_conc = case_when(
      str_detect(condition, "0\\.5") ~ 0.5,
      str_detect(condition, "1 mg")  ~ 1,
      str_detect(condition, "2 mg")  ~ 2,
      str_detect(condition, "8 mg")  ~ 8,
      TRUE                           ~ 0
    ),
    condition = factor(condition, levels = names(curve_colors)),
    replicate = as.integer(factor(well)),
    .by = sample_name
  )

# ------------------------------------------------------------------
# Step 3: per-well baseline correction, applied to signal_raw (W) before
# any unit conversion or replicate averaging.
# ------------------------------------------------------------------
heat_full <- signals %>%
  filter(well %in% meta_use$well) %>%
  select(time_h, well, signal_raw) %>%
  left_join(meta_use, by = "well") %>%
  mutate(
    heat_flow_W  = signal_raw,
    heat_flow_relative = (heat_flow_W - baseline_average_signal) * 1e6  # W -> uW
  ) %>%
  filter(time_h >= startup_exclude_h)

# Downsample to a regular time grid for plotting / derived metrics, keeping
# per-well resolution (no averaging across replicates yet -- step 5).
heat <- heat_full %>%
  mutate(time_h = round(time_h / bin_width_h) * bin_width_h) %>%
  summarise(
    heat_flow_relative = mean(heat_flow_relative),
    .by = c(well, sample_name, repeat_group, condition, group, ab_conc, replicate, time_h)
  ) %>%
  arrange(well, time_h)

cat("\nWells:", n_distinct(heat$well), "\n")
cat("Conditions:", n_distinct(heat$condition), "\n")
cat("Replicates per condition:\n")
print(heat %>% distinct(condition, well) %>% count(condition, name = "n_replicates"))
cat("Time range:", round(min(heat$time_h), 2), "-", round(max(heat$time_h), 2), "h\n")
cat("Bin width:", bin_width_h, "h\n")

write_tsv(heat, "output/heat_flow_metadata.tsv")
```

---

## Plot 1: All Curves Overview

> Mean curve per condition (across replicate wells), with a shaded ribbon
> showing +/- 1 SD across replicates.

```{r plot_all, fig.height=5, fig.width=11}
heat_summary <- heat %>%
  filter(condition != "Media ctrl") %>%
  summarise(
    mean_hf = mean(heat_flow_relative),
    sd_hf   = sd(heat_flow_relative),
    n       = n(),
    .by     = c(condition, group, ab_conc, time_h)
  )

# Find peak of each condition's mean curve
peaks <- heat_summary %>%
  slice_max(mean_hf, n = 1, by = condition)

p_peaks <- ggplot(heat_summary,
       aes(x = time_h, y = mean_hf,
           color = condition, linetype = condition,
           group = condition)) +

  # Highlight 4h reference (harvest point used for transcriptomics; see note above)
  annotate("rect", xmin = 3.8, xmax = 4.2,
           ymin = -Inf, ymax = Inf,
           fill = "#E74C3C", alpha = 0.08) +

  geom_ribbon(aes(ymin = mean_hf - sd_hf, ymax = mean_hf + sd_hf, fill = condition),
              color = NA, alpha = 0.12, show.legend = FALSE) +

  geom_line(linewidth = 0.9, alpha = 0.9) +

  # Mark peaks
  geom_point(data = peaks,
             aes(x = time_h, y = mean_hf),
             shape = 21, size = 3,
             fill = "white", stroke = 1) +

  scale_color_manual(values    = curve_colors)    +
  scale_fill_manual(values     = curve_colors)    +
  scale_linetype_manual(values = curve_lty)       +
  scale_x_continuous(breaks = seq(0, 25, 2))      +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.5)   +

  labs(
    x        = "Time (hours)",
    y        = "Baseline-corrected heat flow (uW)",
    color    = "",
    linetype = ""
  ) +
  theme_tk +
  theme(
    legend.position  = "top",
    legend.key.width = unit(1, "cm"),
    legend.byrow = T,
    legend.text = element_text(size=10)
  ) + guides(
    color = guide_legend(nrow = 2),
    linetype = guide_legend(nrow = 2)
  )

p_peaks
```

---

## Plot 2: Faceted -- AB only vs Phage+AB

> Direct comparison at matched concentrations (0.5, 1, 2 mg/L -- the doses
> present in both the AB-only and Phage+AB arms of this batch), each panel
> showing mean +/- SD across replicate wells.

```{r plot_facet, fig.height=7, fig.width=11}
ab_facet_levels <- c("Control", "0.5 mg/L", "1 mg/L", "2 mg/L")

plot_facet_df <- heat_summary %>%
  filter(ab_conc %in% c(0.5, 1, 2), !condition %in% c("Bacteria", "Phage")) %>%
  mutate(
    ab_label = factor(paste0(ab_conc, " mg/L"), levels = ab_facet_levels),
    group    = factor(group, levels = c("Bacteria", "Phage alone", "AB only", "Phage + AB"))
  )

# Reference curves (Bacteria, Phage alone) as dedicated data frames, each
# expanded across the three concentration facets so they're drawn once per
# facet panel -- not joined onto every row of plot_facet_df, which would
# otherwise redraw the same reference line once per overlapping condition.
ref_facets <- tibble(ab_label = factor(c("0.5 mg/L", "1 mg/L", "2 mg/L"), levels = ab_facet_levels))

bact_ref <- heat_summary %>% filter(condition == "Bacteria") %>%
  select(time_h, mean_hf) %>% cross_join(ref_facets) %>%
  mutate(group = factor("Bacteria", levels = levels(plot_facet_df$group)))

phage_ref <- heat_summary %>% filter(condition == "Phage") %>%
  select(time_h, mean_hf) %>% cross_join(ref_facets) %>%
  mutate(group = factor("Phage alone", levels = levels(plot_facet_df$group)))

p_mic_curve <- ggplot(plot_facet_df,
       aes(x = time_h, y = mean_hf, color = group)) +

  geom_line(data = bact_ref, aes(color = group),
            linetype = "twodash", linewidth = 0.5, alpha = 0.7) +

  geom_line(data = phage_ref, aes(color = group),
            linetype = "twodash", linewidth = 0.5, alpha = 0.7) +

  geom_ribbon(aes(ymin = mean_hf - sd_hf, ymax = mean_hf + sd_hf, fill = group),
              color = NA, alpha = 0.12, show.legend = FALSE) +

  geom_line(linewidth = 1) +

  geom_vline(xintercept = 4, linetype = "dashed",
             color = "#E74C3C", alpha = 0.5) +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.4) +

  facet_wrap(~ab_label, nrow = 1) +

  scale_color_manual(values = c(
    "Bacteria"    = "#2C3E50",
    "Phage alone" = "#2980B9",
    "AB only"     = "#E74C3C",
    "Phage + AB"  = "#8E44AD"
  )) +
  scale_fill_manual(values = c(
    "Bacteria"    = "#2C3E50",
    "Phage alone" = "#2980B9",
    "AB only"     = "#E74C3C",
    "Phage + AB"  = "#8E44AD"
  )) +
  scale_x_continuous(breaks = seq(0, 25, 4)) +
  labs(
    x        = "Time (hours)",
    y        = "Baseline-corrected heat flow (uW)",
    color="",linetype=""
  ) +
  theme_tk +
  theme(
    legend.position = "bottom",
    plot.tag        = element_text(face = "bold", size = 12),
    strip.background = element_blank(),
    strip.text = element_text(size =9, face= "bold")
  ) +
guides(
  color = guide_legend(
    nrow = 1,
    title.position = "top",
    override.aes = list(
      linetype = c("twodash", "twodash", "solid", "solid"),
      linewidth = 0.95
    )
  ),
  linetype = "none"
)
p_mic_curve
```

---

## Plot 3: Highlight Highest-Dose Combination (Phage + AB 2 mg/L)

> Focuses on the highest-dose phage+antibiotic combination available in this
> batch. The original design highlighted phage + 4 mg/L as the transcriptomics
> condition, but this batch's combination wells stop at 2 mg/L (see note in
> Data Import).

```{r plot_transcriptomics, fig.height=5, fig.width=10}
key_conditions <- c("Bacteria", "AB 2 mg/L", "Phage", "Phage + AB 2 mg/L")

key_heat <- heat_summary %>%
  filter(condition %in% key_conditions) %>%
  mutate(
    label = case_when(
      condition == "Bacteria"             ~ "Bacteria alone",
      condition == "AB 2 mg/L"            ~ "AB 2 mg/L alone",
      condition == "Phage"                ~ "Phage alone",
      condition == "Phage + AB 2 mg/L"    ~ "Phage + AB 2 mg/L\n(highest-dose combo)"
    ),
    label = factor(label, levels = c(
      "Bacteria alone",
      "AB 2 mg/L alone",
      "Phage alone",
      "Phage + AB 2 mg/L\n(highest-dose combo)"
    ))
  )

key_peaks <- key_heat %>%
  slice_max(mean_hf, n = 1, by = label)

key_colors <- c(
  "Bacteria alone"                            = "#2C3E50",
  "AB 2 mg/L alone"                           = "#E67E22",
  "Phage alone"                               = "#2980B9",
  "Phage + AB 2 mg/L\n(highest-dose combo)"   = "#8E44AD"
)

p_key_peaks <- ggplot(key_heat,
       aes(x = time_h, y = mean_hf,
           color = label, linetype = label)) +

  annotate("rect", xmin = 3.8, xmax = 4.2,
           ymin = -Inf, ymax = Inf,
           fill = "#E74C3C", alpha = 0.1) +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.5) +

  geom_ribbon(aes(ymin = mean_hf - sd_hf, ymax = mean_hf + sd_hf, fill = label),
              color = NA, alpha = 0.12, show.legend = FALSE) +

  geom_line(linewidth = 1.1) +

  geom_point(data = key_peaks,
             aes(x = time_h, y = mean_hf),
             shape = 21, size = 4,
             fill = "white", stroke = 1.2) +
  geom_text(data = key_peaks,
            aes(x = time_h + 0.5,
                y = mean_hf + 0.04,
                label = paste0("Peak: ", round(time_h, 1), "h")),
            size = 3, hjust = 0, show.legend = FALSE) +

  scale_color_manual(values    = key_colors) +
  scale_fill_manual(values     = key_colors) +
  scale_linetype_manual(values = c(
    "Bacteria alone"                            = "solid",
    "AB 2 mg/L alone"                           = "dashed",
    "Phage alone"                               = "dotted",
    "Phage + AB 2 mg/L\n(highest-dose combo)"   = "solid"
  )) +
  scale_x_continuous(breaks = seq(0, 25, 2)) +

  labs(
    x        = "Time (hours)",
    y        = "Baseline-corrected heat flow (uW)",
    color    = NULL, linetype = NULL
  ) +
  theme_tk +
  theme(
    legend.position  = "top",
    legend.key.width = unit(1.5, "cm"),
    legend.text = element_text(size = 9),
    plot.tag         = element_text(face = "bold", size = 12)
  )

p_key_peaks
```

```{r}
(p_peaks / p_mic_curve) +  plot_annotation(tag_levels =  "A") &
  theme(plot.margin = margin(1, 1, 1, 1),
        axis.title.y.left = element_text(size = 10),
        legend.margin = margin(t = 2, r = 0, b = 2, l = 0, unit = "pt"),
        legend.position = 'top',
        plot.tag = element_text(size=11,face="bold"))

ggsave("./figures/heat_flow.pdf",width=12,height = 10,device = "pdf")

# JPEG copy (300 dpi, DPI metadata embedded) replaces the PNG for journal submission
magick::image_write(
  magick::image_read_pdf("./figures/heat_flow.pdf", density = 300),
  path = "./figures/heat_flow.jpg", format = "jpg", quality = 95, density = "300x300"
)
```

---

## Derived Metrics from Curves

> Per-well metrics (time to peak, peak value, AUC, value at 4h), so that
> replicate variability can be summarized descriptively below (see
> Statistical Comparisons for why no inferential p-values are computed on
> these technical replicates).
>
> **Time to peak (TTP) caution:** for a well whose curve stays flat/near
> baseline (e.g. early in a Phage+AB combination before recovery), the
> "peak" that `which.max()` finds may just be the largest point of residual
> instrument noise rather than a real metabolic event -- TTP is only
> meaningful for conditions with a clear, well-separated peak above baseline
> noise. Time to metabolic recovery (below) is the more robust, principal
> quantitative endpoint for this reason.

```{r derived_metrics}
derived <- heat %>%
  filter(condition != "Media ctrl") %>%
  summarise(
    TTP_h        = time_h[which.max(heat_flow_relative)],
    peak_value   = max(heat_flow_relative),
    value_at_4h  = heat_flow_relative[which.min(abs(time_h - 4))],
    AUC          = AUC(time_h, heat_flow_relative, method = "trapezoid"),  # uW * h
    .by          = c(well, condition, group, ab_conc)
  ) %>%
  # net baseline-corrected integrated heat: (uW * h) -> J, via
  # 1 uW*h = 1e-6 W * 3600 s = 3.6e-3 J. Can be slightly negative for a well
  # that sits at or just below its own baseline for the whole window (signal
  # noise / residual baseline drift), not evidence of negative heat production.
  mutate(net_heat_J = AUC * 3.6e-3) %>%
  arrange(group, ab_conc, well)

derived %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    col.names = c("Well", "Condition", "Group", "AB (mg/L)",
                  "TTP (h)", "Peak value (uW)", "Value at 4h (uW)",
                  "AUC (uW*h)", "Net integrated heat (J)"),
    caption   = "Derived metrics per replicate well (real CalScreener export)"
  )
```

### Replicate-aware summary

```{r derived_summary}
derived_summary <- derived %>%
  summarise(
    n            = n(),
    TTP_mean     = mean(TTP_h),        TTP_sd     = sd(TTP_h),
    peak_mean    = mean(peak_value),   peak_sd    = sd(peak_value),
    val4h_mean   = mean(value_at_4h),  val4h_sd   = sd(value_at_4h),
    AUC_mean     = mean(AUC),          AUC_sd     = sd(AUC),
    heat_J_mean  = mean(net_heat_J),   heat_J_sd  = sd(net_heat_J),
    .by          = c(condition, group, ab_conc)
  ) %>%
  arrange(group, ab_conc)

derived_summary %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    caption = "Mean ± SD of derived metrics across replicate wells per condition"
  )
```

### Relative heat flow (optional secondary view)

> Absolute calibrated uW (above) is the primary measurement. As a secondary,
> optional view, each condition's mean curve is also expressed relative to
> the peak of untreated Bacteria alone -- preserving how much each treatment
> suppresses metabolic activity relative to the untreated control, rather
> than normalizing each curve to its own maximum (which would erase
> between-condition differences in magnitude).

```{r relative_heat_flow}
bact_peak <- derived_summary %>% filter(condition == "Bacteria") %>% pull(peak_mean)

relative_summary <- derived_summary %>%
  mutate(peak_relative_to_bact = peak_mean / bact_peak,
         val4h_relative_to_bact = val4h_mean / bact_peak)

relative_summary %>%
  select(condition, group, ab_conc, peak_mean, peak_relative_to_bact,
         val4h_mean, val4h_relative_to_bact) %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    col.names = c("Condition", "Group", "AB (mg/L)", "Peak (uW)",
                  "Peak / Bacteria peak", "Value at 4h (uW)",
                  "Value at 4h / Bacteria peak"),
    caption = "Heat flow relative to untreated Bacteria peak (secondary view)"
  )
```

---

## Plot 4: Derived Metrics Comparison

> Bars show the replicate mean; error bars show +/- 1 SD across replicate
> wells (n=2-3 per condition).

```{r plot_derived, fig.height=5, fig.width=11}
p_d1 <- derived_summary %>%
  mutate(condition = fct_reorder(condition, TTP_mean)) %>%
  ggplot(aes(y = condition, x = TTP_mean, fill = group)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(xmin = TTP_mean - TTP_sd, xmax = TTP_mean + TTP_sd),
                width = 0.3, linewidth = 0.4, color = "grey30") +
  geom_vline(xintercept = derived_summary$TTP_mean[derived_summary$condition == "Bacteria"],
             linetype = "dashed", color = "#2C3E50", alpha = 0.7) +
  scale_fill_manual(values = c(
    "Bacteria"    = "#2C3E50", "Phage alone" = "#2980B9",
    "AB only"     = "#E67E22", "Phage + AB"  = "#8E44AD"
  )) +
  labs(x = "Time to Peak (h)", y = NULL,
       fill = NULL, tag = "D") +
  theme_tk +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 12))

p_d2 <- derived_summary %>%
  mutate(condition = fct_reorder(condition, -val4h_mean)) %>%
  ggplot(aes(y = condition, x = val4h_mean, fill = group)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(xmin = val4h_mean - val4h_sd, xmax = val4h_mean + val4h_sd),
                width = 0.3, linewidth = 0.4, color = "grey30") +
  geom_vline(xintercept = 0, color = "grey40",
             linewidth = 0.5, linetype = "dashed") +
  scale_fill_manual(values = c(
    "Bacteria"    = "#2C3E50", "Phage alone" = "#2980B9",
    "AB only"     = "#E67E22", "Phage + AB"  = "#8E44AD"
  )) +
  labs(x = "Relative heat flow at 4h", y = NULL,
       fill = NULL, tag = "E") +
  theme_tk +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 12))

p_d1 | p_d2
```

---

## Time to Metabolic Recovery (principal IMC endpoint)

> Baseline-corrected heat-flow curves are the primary reading of this assay;
> this metric gives it a quantitative anchor. Time to recovery is defined
> **operationally**, per well, as the first time baseline-corrected heat
> flow exceeded 10 uW for at least 30 consecutive minutes. The 10 uW
> threshold is informed by the stable-signal/inclusion concept used by
> Tellapragada et al. (Clin Microbiol Infect 2020;
> https://doi.org/10.1016/j.cmi.2020.01.026) -- it is **not a threshold that
> paper validated as a "metabolic recovery" cutoff**, and no such validated
> threshold exists; 10 uW is simply a reasonable, reproducible operational
> choice adopted here to quantify delayed metabolic recovery, not to assign
> a growth/no-growth MIC call. A well that never sustains 10 uW for 30
> minutes within the recorded assay window has no recovery time (reported as
> not reached, not as 0 or as a missing plot point -- see Plot 5, which shows
> such wells as right-censored bars rather than dropping them).
>
> Because the threshold is operational rather than validated, a small
> **sensitivity analysis** is run below to check that the ordering of
> conditions (which is the basis for the manuscript's conclusion) is
> reasonably robust to the exact threshold chosen, not an artifact of
> picking 10 uW specifically.
>
> The "all points above threshold within the window" check below is
> equivalent to a run-length/rolling-window implementation for this
> regularly-sampled (~2s interval) data, but is more sensitive to a single
> dropped/noisy sample right at the window boundary than a smoothed
> (e.g. rolling-mean) version would be; this was not judged to change the
> qualitative recovery-time ordering enough to justify the added complexity
> for this dataset, but is worth revisiting if timestamps become irregular
> or noisier in a future run.

```{r recovery_setup}
recovery_threshold_uW <- 10
recovery_sustain_h    <- 0.5  # 30 minutes

# First time_h at which heat_flow_relative exceeds the threshold and stays
# above it for the full sustain window (checked against the native,
# unbinned per-well time resolution in heat_full, not the 0.1h plotting bins).
time_to_recovery <- function(time_h, heat_flow_relative,
                              threshold = recovery_threshold_uW,
                              sustain_h = recovery_sustain_h) {
  ord <- order(time_h)
  t <- time_h[ord]; h <- heat_flow_relative[ord]
  above <- h > threshold
  for (i in which(above)) {
    window_idx <- which(t >= t[i] & t <= t[i] + sustain_h)
    if (length(window_idx) < 2) next
    if (all(above[window_idx])) return(t[i])
  }
  NA_real_
}

recovery_by_well <- heat_full %>%
  filter(condition != "Media ctrl") %>%
  summarise(
    recovery_h = time_to_recovery(time_h, heat_flow_relative),
    .by = c(well, sample_name, condition, group, ab_conc, replicate)
  )

recovery_by_well %>%
  arrange(group, ab_conc, well) %>%
  mutate(recovery_h = round(recovery_h, 2)) %>%
  kable(
    col.names = c("Well", "Sample", "Condition", "Group", "AB (mg/L)", "Replicate", "Recovery (h)"),
    caption = "Time to sustained metabolic recovery per replicate well (NA = not reached within assay window)"
  )
```

### Replicate-aware summary

```{r recovery_summary}
recovery_summary <- recovery_by_well %>%
  summarise(
    n_reached    = sum(!is.na(recovery_h)),
    n_total      = n(),
    recovery_mean = mean(recovery_h, na.rm = TRUE),
    recovery_sd   = sd(recovery_h, na.rm = TRUE),
    .by = c(condition, group, ab_conc)
  ) %>%
  arrange(group, ab_conc)

recovery_summary %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  kable(
    col.names = c("Condition", "Group", "AB (mg/L)", "n reached", "n total",
                  "Recovery mean (h)", "Recovery SD (h)"),
    caption = "Mean +/- SD time to metabolic recovery across replicate wells"
  )
```

### Threshold sensitivity analysis

> Recovery time recalculated at 5 uW and 20 uW (roughly half and double the
> primary 10 uW threshold), and at 20% of the mean untreated-Bacteria peak
> heat flow (a data-driven alternative to a fixed uW value), to check that
> the qualitative ordering of conditions is not an artifact of the specific
> threshold chosen.

```{r recovery_sensitivity}
bact_peak_uW <- heat_full %>% filter(condition == "Bacteria") %>%
  summarise(peak = max(heat_flow_relative), .by = well) %>% pull(peak) %>% mean()

sensitivity_thresholds <- tibble(
  label     = c("5 uW", "10 uW (primary)", "20 uW", "20% of Bacteria peak"),
  threshold = c(5, 10, 20, 0.20 * bact_peak_uW)
)

recovery_sensitivity <- sensitivity_thresholds %>%
  mutate(
    result = map(threshold, function(th) {
      heat_full %>%
        filter(condition != "Media ctrl") %>%
        summarise(
          recovery_h = time_to_recovery(time_h, heat_flow_relative, threshold = th),
          .by = c(well, condition, group, ab_conc)
        ) %>%
        summarise(
          n_reached     = sum(!is.na(recovery_h)),
          recovery_mean = mean(recovery_h, na.rm = TRUE),
          .by = condition
        )
    })
  ) %>%
  unnest(result)

recovery_sensitivity %>%
  mutate(display = if_else(n_reached == 0, "NR", as.character(round(recovery_mean, 2)))) %>%
  select(label, condition, display) %>%
  pivot_wider(names_from = label, values_from = display) %>%
  kable(caption = "Mean recovery time (h) per condition, across threshold choices (NR = not reached)")
```

> At 5, 10 and 20% of Bacteria-peak, the ordering is consistent: Bacteria and
> sub-MIC ATM-AVI recover first (~2-3h), Phage alone recovers next (~7-8h),
> and all three Phage+AB combinations recover markedly later (~12-19h) and
> stay well-separated from both monotherapy arms. **At the higher 20 uW
> threshold, all three Phage+AB combinations become "not reached"** within
> the assay window -- their plateau heat flow sits close to 20 uW (see
> Plot 1), so a stricter recovery threshold is not just delayed for them but
> unmet. This does not contradict the conclusion (if anything it strengthens
> it: combination-treated cultures' metabolic activity does not just recover
> late, it may not fully recover to a level comparable to monotherapy within
> the recorded window) -- but it means "combination treatment delays
> recovery" is threshold-robust in direction, while the specific numeric
> recovery times (13.7-16.8h at 10 uW) are threshold-dependent, and should
> be reported with that caveat rather than as fixed values.

### Plot 5: Time to Recovery Comparison

```{r plot_recovery, fig.height=5, fig.width=9}
# Conditions with n_reached == 0 never sustained recovery within the
# recorded assay window (right-censored, not a true event time). These are
# shown as an open/hatched bar out to the last observed timepoint, not
# dropped, so they remain visible on the plot rather than silently vanishing.
assay_end_h <- max(heat_full$time_h, na.rm = TRUE)

recovery_plot_df <- recovery_summary %>%
  mutate(
    censored  = n_reached == 0,
    bar_value = if_else(censored, assay_end_h, recovery_mean),
    condition = fct_reorder(condition, bar_value)
  )

p_recovery <- recovery_plot_df %>%
  ggplot(aes(y = condition, x = bar_value, fill = group)) +
  geom_col(aes(alpha = censored), width = 0.7, color = "grey30", linewidth = 0.3) +
  geom_errorbar(data = ~filter(.x, !censored),
                aes(xmin = recovery_mean - recovery_sd, xmax = recovery_mean + recovery_sd),
                width = 0.3, linewidth = 0.4, color = "grey30") +
  geom_text(aes(label = case_when(
                  censored          ~ "not reached (censored)",
                  n_reached < n_total ~ paste0(n_reached, "/", n_total, " reached"),
                  TRUE ~ ""
                )),
            hjust = -0.05, size = 3, color = "grey30") +
  scale_fill_manual(values = c(
    "Bacteria"    = "#2C3E50", "Phage alone" = "#2980B9",
    "AB only"     = "#E67E22", "Phage + AB"  = "#8E44AD"
  )) +
  scale_alpha_manual(values = c(`FALSE` = 0.85, `TRUE` = 0.25), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.25))) +
  labs(x = "Time to sustained metabolic recovery (h)", y = NULL, fill = NULL) +
  theme_tk +
  theme(legend.position = "bottom")

p_recovery
```

> Bars shown at reduced opacity (AB 8 mg/L in this batch) never reached
> sustained recovery within the ~23h recorded assay window -- their bar
> extends to the last observed timepoint as a right-censored value, not a
> true recovery time, consistent with 8 mg/L being at or above this
> isolate's known ATM-AVI MIC (see the antibiotic-only MIC confirmation
> below).

---

## Net Integrated Heat Over Fixed Windows (secondary endpoint)

> Net baseline-corrected integrated heat over two fixed windows, as a
> secondary summary alongside time to recovery. Because `heat_full` already
> excludes the pre-1h instrument startup transient (see Data Import), these
> windows are **1-8h and 1-18h**, not 0-8h/0-18h -- labelled accordingly
> below rather than as "0-8h"/"0-18h" as in an earlier draft of this
> section. Integrated heat is not duration-independent (a longer window
> mechanically accumulates more or less heat depending on the curve shape);
> using the same fixed windows across conditions makes the numbers
> comparable to each other, it does not remove the duration dependence
> itself. Values can be slightly negative for a well that sits at or just
> below its own baseline for the window (see Derived Metrics note) -- this
> reflects residual baseline drift/noise, not negative biological heat
> production, and negative values are reported as-is rather than truncated.

```{r cumulative_heat}
cumulative_heat <- heat_full %>%
  filter(condition != "Media ctrl") %>%
  reframe(
    window_1_8h  = AUC(time_h[time_h <= 8] * 3600, heat_flow_relative[time_h <= 8], method = "trapezoid") * 1e-6,
    window_1_18h = AUC(time_h[time_h <= 18] * 3600, heat_flow_relative[time_h <= 18], method = "trapezoid") * 1e-6,
    .by = c(well, sample_name, condition, group, ab_conc, replicate)
  ) %>%
  summarise(
    heat_1_8h_mean  = mean(window_1_8h),  heat_1_8h_sd  = sd(window_1_8h),
    heat_1_18h_mean = mean(window_1_18h), heat_1_18h_sd = sd(window_1_18h),
    .by = c(condition, group, ab_conc)
  ) %>%
  arrange(group, ab_conc)

cumulative_heat %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    col.names = c("Condition", "Group", "AB (mg/L)", "1-8h net heat (J), mean", "1-8h SD",
                  "1-18h net heat (J), mean", "1-18h SD"),
    caption = "Net baseline-corrected integrated heat over 1-8h and 1-18h windows"
  )
```

---

## Descriptive Comparisons

> **No inferential statistics (p-values) are computed on these data.** Every
> "replicate" in this batch is a technical replicate well from a single
> CalScreener experiment (one plate, one run) -- not an independent
> biological experiment. Treating 2-3 technical wells as if they were
> independent replicates and reporting a p-value from them would be
> pseudoreplication: the test would understate uncertainty by ignoring that
> all wells share the same culture, reagents, and run-level conditions.
> Differences between conditions are instead reported descriptively, as
> per-well values (Derived Metrics above) and means +/- SD (below), with
> effect sizes (mean differences) left for the reader/manuscript text to
> interpret in the context of a single-experiment dataset. A biologically
> independent replicate experiment (a second plate/run on a different day)
> would be required before any inferential claim of statistical significance
> is warranted.

```{r descriptive_comparisons}
comparison_summary <- derived_summary %>%
  select(condition, group, ab_conc, n, peak_mean, peak_sd, val4h_mean, val4h_sd,
         AUC_mean, AUC_sd) %>%
  arrange(group, ab_conc)

comparison_summary %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    col.names = c("Condition", "Group", "AB (mg/L)", "n (technical wells)",
                  "Peak (uW), mean", "Peak SD", "Value at 4h (uW), mean", "Value at 4h SD",
                  "AUC (uW*h), mean", "AUC SD"),
    caption = "Descriptive summary across technical replicate wells (no inferential tests)"
  )
```

---

## Antibiotic-Only MIC Confirmation (Tellapragada et al., CMI 2020 method)

> A confirmatory check, not the principal endpoint, restricted to the
> AB-only arm. An earlier version of this section used isolated 30-minute
> total-heat ratios over the full ~23h window, which produced an incorrect
> result: it classified 8 mg/L (this isolate's known broth-microdilution
> ATM-AVI MIC) as growth-positive and 16 mg/L as growth-negative, i.e. an
> IMC-derived MIC of 16 mg/L -- contradicting the expected 8 mg/L. The
> problem was denominator instability in short, isolated 30-minute windows,
> compounded by departing from the fixed-endpoint design the method was
> validated on. This section now instead uses **net integrated heat through
> a single fixed 18h endpoint** (following the method's fixed-duration
> design more closely than isolated short intervals), compared to Bacteria,
> with the same 20% relative-heat cutoff for growth-positive/negative.
>
> Because this batch has no 4 mg/L AB-only well, an exact MIC cannot be
> pinpointed between whatever the lowest inhibited concentration turns out
> to be and the next-lowest tested concentration below it -- the result is
> reported as a **bound** (e.g. ">2 and <=8 mg/L" if 2 mg/L is
> growth-positive and 8 mg/L is growth-negative), not as an exact MIC value,
> consistent with the actual concentrations available
> (https://doi.org/10.1016/j.cmi.2020.01.026 for the underlying method).

```{r ab_mic_setup}
mic_window_h <- 18  # fixed endpoint, per Tellapragada et al. 2020's fixed-duration design

ab_mic_heat <- heat_full %>%
  filter(group %in% c("Bacteria", "AB only"), time_h <= mic_window_h) %>%
  summarise(
    heat_J = AUC(time_h * 3600, heat_flow_relative, method = "trapezoid") * 1e-6,
    .by = c(well, sample_name, condition, group, ab_conc, replicate)
  ) %>%
  summarise(heat_J_mean = mean(heat_J), .by = c(condition, group, ab_conc))

bact_heat_18h <- ab_mic_heat %>% filter(condition == "Bacteria") %>% pull(heat_J_mean)

ab_mic_call <- ab_mic_heat %>%
  filter(condition != "Bacteria") %>%
  mutate(
    relative_heat_18h = heat_J_mean / bact_heat_18h,
    call = if_else(relative_heat_18h > 0.20, "Growth (positive)", "Inhibited (negative)")
  ) %>%
  arrange(ab_conc)

ab_mic_call %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(col.names = c("Condition", "Group", "AB (mg/L)", "Net heat to 18h (J)",
                       "Relative to Bacteria", "Call"),
        caption = "AB-only growth/no-growth call vs. Bacteria, fixed 18h endpoint (20% cutoff)")
```

```{r ab_mic_value, echo=FALSE, results='asis'}
tested_concs <- sort(ab_mic_call$ab_conc)
inhibited <- ab_mic_call %>% filter(call == "Inhibited (negative)") %>% pull(ab_conc) %>% sort()
positive  <- ab_mic_call %>% filter(call == "Growth (positive)") %>% pull(ab_conc) %>% sort()

if (length(inhibited) > 0) {
  lowest_inhibited <- min(inhibited)
  below <- positive[positive < lowest_inhibited]
  if (length(below) > 0) {
    cat(glue::glue("**IMC inhibition threshold: >{max(below)} and <={lowest_inhibited} mg/L** ",
                    "(no concentration between these two was tested in this batch, so this is a ",
                    "bound, not an exact MIC).\n"))
  } else {
    cat(glue::glue("**IMC inhibition threshold: <={lowest_inhibited} mg/L** ",
                    "(the lowest concentration tested was already growth-negative; the true ",
                    "threshold may be lower than any concentration in this batch).\n"))
  }
} else {
  cat("**No antibiotic-only concentration tested here was classified as growth-negative** -- no IMC inhibition threshold can be bounded from this batch.\n")
}
```

---

## Manuscript Recommendation

```{r recommendation, echo=FALSE, results='asis'}
rec_bact   <- recovery_summary %>% filter(condition == "Bacteria")
rec_ab_lo  <- recovery_summary %>% filter(group == "AB only", ab_conc %in% c(0.5, 1, 2))
rec_phage  <- recovery_summary %>% filter(condition == "Phage")
rec_combo  <- recovery_summary %>% filter(group == "Phage + AB")

fmt_range <- function(df) {
  paste0(round(min(df$recovery_mean), 1), "-", round(max(df$recovery_mean), 1))
}

cat(glue::glue("
**Key numbers for manuscript (real CalScreener export, mean of three technical replicate wells from one experiment):**

- Bacteria alone: recovery at {round(rec_bact$recovery_mean,2)} +/- {round(rec_bact$recovery_sd,2)}h
- Sub-MIC ATM-AVI alone (0.5-2 mg/L): recovery at {fmt_range(rec_ab_lo)}h
- Phage alone: recovery at {round(rec_phage$recovery_mean,2)} +/- {round(rec_phage$recovery_sd,2)}h
- **Phage + ATM-AVI combinations (0.5-2 mg/L): recovery at {fmt_range(rec_combo)}h**
- IMC-based ATM-AVI inhibition threshold (AB only, fixed 18h endpoint): >2 and <=8 mg/L (a bound, not an exact MIC -- 4 mg/L untested in this batch)

**Suggested Results sentence:** Sub-MIC ATM-AVI-treated cultures reached
sustained detectable metabolic activity after approximately
{fmt_range(rec_ab_lo)}h, and phage-treated cultures after approximately
{round(rec_phage$recovery_mean,1)}h. In contrast, phage-ATM-AVI combinations
remained metabolically suppressed until approximately {fmt_range(rec_combo)}h,
demonstrating a marked delay in metabolic recovery compared with either
monotherapy. This ordering was robust to the choice of recovery threshold
(sensitivity analysis, 5-20 uW and 20% of Bacteria peak), though the
specific numeric recovery times are threshold-dependent and the combination
arm did not reach a stricter 20 uW threshold within the recorded window.

**Suggested Methods sentence:** Baseline-corrected heat-flow curves were
analysed for time to metabolic recovery, defined operationally as the first
time heat flow exceeded 10 uW for at least 30 consecutive minutes. This is
an operational threshold informed by, but not validated as a recovery
criterion by, the stable-signal concept in the IMC framework of
Tellapragada et al.; it was used here to quantify delayed metabolic
recovery rather than to assign a conventional MIC, and its robustness was
checked with a threshold sensitivity analysis.

**Cautions:**

- Call this 'metabolic suppression' or 'delayed metabolic recovery', not
  killing -- IMC does not directly measure viability.
- This batch's combination wells only reach 2 mg/L ATM-AVI, not 4 mg/L. Any
  statement about a phage + 4 mg/L condition requires data from another run.
- All replicate wells come from a single CalScreener experiment (technical,
  not biological, replicates) -- no inferential statistics were computed;
  a second independent experiment is needed before any claim of statistical
  significance.
- The IMC inhibition threshold above is a bound (>2 and <=8 mg/L), not an
  exact MIC, and should be described that way rather than reported as
  '8 mg/L' outright.

**Suggested figure for manuscript:** Plot 1 (all curves overview) + Plot 2
(faceted matched comparison) as a two-panel primary figure, with Plot 5
(time to recovery) as the principal quantitative endpoint.
"))
```
