---
title: "IMC Heat Flow Curve Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 150)
```

```{r libraries}
rm(list=ls())
library(tidyverse)
library(patchwork)
library(knitr)

theme_tk <- theme_bw(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "#F0F0F0"),
    plot.title       = element_text(face = "bold", size = 13)
  )

# ============================================================
# USER PARAMETERS
# ============================================================

heat_file <- "../../time_kill_data/heat_flow_digitized_0.25h_from_heat_png.csv"

# Colors per condition
curve_colors <- c(
  "Bacteria"       = "#2C3E50",   # Dark grey
  "AB 8 mg/L"      = "#922B21",   # Dark red
  "AB 4 mg/L"      = "#E74C3C",   # Red
  "AB 2 mg/L"      = "#E67E22",   # Orange
  "AB 1 mg/L"      = "#F39C12",   # Amber
  "AB 0.5 mg/L"    = "#FAD7A0",   # Light amber
  "Phage"    = "#2980B9",   # Blue
  "Media ctrl"     = "#BDC3C7",   # Grey
  "Phage + AB 4 mg/L" = "#6C3483",   # Dark purple  ← transcriptomics condition
  "Phage + AB 2 mg/L" = "#8E44AD",   # Purple
  "Phage + AB 1 mg/L" = "#BB8FCE",   # Medium purple
  "Phage + AB 0.5 mg/L"= "#D7BDE2"  # Light purple
)

# Line types
curve_lty <- c(
  "Bacteria"        = "solid",
  "AB 8 mg/L"       = "dashed",
  "AB 4 mg/L"       = "dashed",
  "AB 2 mg/L"       = "dashed",
  "AB 1 mg/L"       = "dashed",
  "AB 0.5 mg/L"     = "dashed",
  "Phage"     = "dotdash",
  "Media ctrl"      = "dotted",
  "Phage + AB 4 mg/L"  = "solid",    # Highlight
  "Phage + AB 2 mg/L"  = "solid",
  "Phage + AB 1 mg/L"  = "solid",
  "Phage + AB 0.5 mg/L"= "solid"
)
```

## Data Import & Cleaning

```{r data_prep}
raw <- read.csv(heat_file, stringsAsFactors = FALSE)

heat <- raw %>%
  mutate(
    condition = case_when(
      well == "B1" ~ "Bacteria",
      well == "B2" ~ "AB 8 mg/L",
      well == "B3" ~ "AB 4 mg/L",
      well == "B4" ~ "AB 2 mg/L",
      well == "B5" ~ "AB 1 mg/L",
      well == "B6" ~ "AB 0.5 mg/L",
      well == "C1" ~ "Phage",
      well == "C2" ~ "Media ctrl",
      well == "C3" ~ "Phage + AB 4 mg/L",
      well == "C4" ~ "Phage + AB 2 mg/L",
      well == "C5" ~ "Phage + AB 1 mg/L",
      well == "C6" ~ "Phage + AB 0.5 mg/L"
    ),
    group = case_when(
      well %in% c("B2","B3","B4","B5","B6") ~ "AB only",
      well %in% c("C3","C4","C5","C6")      ~ "Phage + AB",
      well == "C1"                           ~ "Phage + bacteria",
      well == "B1"                           ~ "Bacteria",
      TRUE                                   ~ "Control"
    ),
    # Extract AB concentration
    ab_conc = case_when(
      str_detect(condition, "0\\.5") ~ 0.5,
      str_detect(condition, "1 mg")  ~ 1,
      str_detect(condition, "2 mg")  ~ 2,
      str_detect(condition, "4 mg")  ~ 4,
      str_detect(condition, "8 mg")  ~ 8,
      TRUE                           ~ 0
    ),
    condition = factor(condition, levels = names(curve_colors)),
    # Flag transcriptomics condition
    is_transcriptomics = well == "C3"
  )

cat("Wells:", unique(heat$well), "\n")
cat("Time range:", min(heat$time_h), "–", max(heat$time_h), "h\n")
cat("Resolution: 0.25h\n")
cat("Transcriptomics condition present (C3 = Phage + 4mg/L):", 
    "C3" %in% unique(heat$well), "\n")

write_tsv(heat,'../../time_kill_data/heat_flow_metadata.tsv')
```

---

## Plot 1: All Curves Overview

> Full reproduction of Figure 2 from the digitized data.

```{r plot_all, fig.height=5, fig.width=11}
# Find peak of each curve
peaks <- heat %>%
  filter(!well %in% c("C2")) %>%   # Exclude media control
  group_by(condition, well) %>%
  slice_max(heat_flow_relative, n = 1) %>%
  ungroup()

p_peaks <- ggplot(heat %>% filter(well != "C2"),
       aes(x = time_h, y = heat_flow_relative,
           color = condition, linetype = condition,
           group = condition)) +

  # # Highlight transcriptomics timepoint
  annotate("rect", xmin = 3.8, xmax = 4.2,
           ymin = -Inf, ymax = Inf,
           fill = "#E74C3C", alpha = 0.08) +
  # annotate("text", x = 4, y = max(heat$heat_flow_relative) * 1.02,
  #          label = "4h\n(Transcriptomics)", size = 2.8,
  #          color = "#E74C3C", fontface = "italic", hjust = 0.5) +

  geom_line(linewidth = 0.9, alpha = 0.9) +

  # Mark peaks
  geom_point(data = peaks,
             aes(x = time_h, y = heat_flow_relative),
             shape = 21, size = 3,
             fill = "white", stroke = 1) +

  scale_color_manual(values    = curve_colors)    +
  scale_linetype_manual(values = curve_lty)       +
  scale_x_continuous(breaks = seq(0, 25, 2))      +
  scale_y_continuous(breaks = seq(-1, 2, 0.25))   +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.5)   +

  labs(
    #title    = "IMC Relative Heat Flow — All Conditions",
    #subtitle = "Open circles = peak metabolic activity | Red band = transcriptomics harvest point (4h)",
    x        = "Time (hours)",
    y        = "Relative heat flow",
    color    = "",
    linetype = ""
  ) +
  theme_tk +
  theme(
    legend.position  = "top",
    legend.key.width = unit(1, "cm"),
    legend.byrow = T,
    legend.text = element_text(size=10)
  ) + guides(
    color = guide_legend(nrow = 2),
    linetype = guide_legend(nrow = 2)
  )

p_peaks
```

---

## Plot 2: Faceted — AB only vs Phage+AB

> Direct comparison at matched concentrations.

```{r plot_facet, fig.height=7, fig.width=11}
# Only matched concentrations (0.5, 1, 2, 4 mg/L)
matched_heat <- heat %>%
  filter(ab_conc %in% c(0.5, 1, 2, 4) | well %in% c("B1","C1")) %>%
  filter(well != "C2") %>%
  mutate(
    ab_label  = ifelse(ab_conc == 0, "Control", paste0(ab_conc, " mg/L")),
    ab_label  = factor(ab_label,
                       levels = c("Control", "0.5 mg/L",
                                  "1 mg/L", "2 mg/L", "4 mg/L"))
  )

# Reference lines: bacteria and phage alone
bact_curve  <- heat %>% filter(well == "B1") %>%
  select(time_h, heat_flow_relative) %>%
  rename(bact_heat = heat_flow_relative)

phage_curve <- heat %>% filter(well == "C1") %>%
  select(time_h, heat_flow_relative) %>%
  rename(phage_heat = heat_flow_relative)

plot_facet_df <- matched_heat %>%
  filter(!well %in% c("B1", "C1")) %>%
  left_join(bact_curve,  by = "time_h") %>%
  left_join(phage_curve, by = "time_h")

# Create dummy rows for the legend
dummy_legend <- tibble(
  time_h = 1,
  heat_flow_relative = 0,
  group = c("Bacteria", "Phage"),
  ab_label = c("0.5 mg/L","0.5 mg/L")
)
df <- bind_rows(plot_facet_df,dummy_legend)
df <- df %>% mutate(group=factor(group,levels=c("Bacteria","Phage","AB only","Phage + AB")))

p_mic_curve <- ggplot(df,
       aes(x = time_h, y = heat_flow_relative,
           color = group)) +

  # Bacteria reference (grey background line)
  geom_line(aes(y = bact_heat),
            color = "#2C3E50", linetype = "twodash",
            linewidth = 0.5, alpha = 0.7) +

  # Phage alone reference (blue background line)
  geom_line(aes(y = phage_heat),
            color = "#2980B9", linetype = "twodash",
            linewidth = 0.5, alpha = 0.7) +

  # Main curves
  geom_line(linewidth = 1) +

  # Highlight 4h
  geom_vline(xintercept = 4, linetype = "dashed",
             color = "#E74C3C", alpha = 0.5) +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.4) +

  facet_wrap(~ab_label, nrow = 1) +

  scale_color_manual(values = c(
    "Bacteria" = "#2C3E50",
    "Phage" = "#2980B9",
    "AB only"    = "#E74C3C",
    "Phage + AB" = "#8E44AD"
  )) +
  scale_linetype_manual(values = c(
    "Bacteria" = "twodash",
    "Phage" = "twodash",
    "AB only"    = "solid",
    "Phage + AB" = "solid"
  )) +
  scale_x_continuous(breaks = seq(0, 25, 4)) +
  labs(
    #title    = "IMC Curves: AB Only vs Phage + AB at Matched Concentrations",
    #subtitle = "Grey = bacteria alone | Blue dotted = phage alone | Red line = 4h transcriptomics point",
    x        = "Time (hours)",
    y        = "Relative heat flow",
    color="",linetype=""
  ) +
  theme_tk +
  theme(
    legend.position = "bottom",
    plot.tag        = element_text(face = "bold", size = 12),
    strip.background = element_blank(),
    strip.text = element_text(size =9, face= "bold")
  ) + 
guides(
  color = guide_legend(
    nrow = 1,
    title.position = "top",
    override.aes = list(
      linetype = c("twodash", "twodash", "solid", "solid"),
      linewidth = 0.95
    )
  ),
  linetype = "none"      # Hide duplicate linetype legend
)
p_mic_curve
```

---

## Plot 3: Highlight Transcriptomics Condition (C3 = Phage + 4 mg/L)

> Focuses on the specific condition used for transcriptomics.

```{r plot_transcriptomics, fig.height=5, fig.width=10}
# Key curves for this comparison
key_wells <- c("B1", "B3", "C1", "C3")   # Bacteria, AB 4mg/L, Phage, Phage+4mg/L

key_heat <- heat %>%
  filter(well %in% key_wells) %>%
  mutate(
    label = case_when(
      well == "B1" ~ "Bacteria alone",
      well == "B3" ~ "AB 4 mg/L alone",
      well == "C1" ~ "Phage alone",
      well == "C3" ~ "Phage + AB 4 mg/L\n(Transcriptomics condition)"
    ),
    label = factor(label, levels = c(
      "Bacteria alone",
      "AB 4 mg/L alone",
      "Phage alone",
      "Phage + AB 4 mg/L\n(Transcriptomics condition)"
    ))
  )

# Peak values for annotation
key_peaks <- key_heat %>%
  group_by(label, well) %>%
  slice_max(heat_flow_relative, n = 1) %>%
  ungroup()

key_colors <- c(
  "Bacteria alone"                              = "#2C3E50",
  "AB 4 mg/L alone"                             = "#E74C3C",
  "Phage alone"                                 = "#2980B9",
  "Phage + AB 4 mg/L" = "#6C3483"
)

p_key_peaks <- ggplot(key_heat,
       aes(x = time_h, y = heat_flow_relative,
           color = label, linetype = label)) +

  # 4h harvest shading
  annotate("rect", xmin = 3.8, xmax = 4.2,
           ymin = -Inf, ymax = Inf,
           fill = "#E74C3C", alpha = 0.1) +
  # annotate("text", x = 4.3, y = 0.95,
  #          label = "4h harvest", size = 3,
  #          color = "#E74C3C", fontface = "bold", hjust = 0) +

  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.5) +
  geom_line(linewidth = 1.1) +

  # Peak markers
  geom_point(data = key_peaks,
             aes(x = time_h, y = heat_flow_relative),
             shape = 21, size = 4,
             fill = "white", stroke = 1.2) +
  geom_text(data = key_peaks,
            aes(x = time_h + 0.5,
                y = heat_flow_relative + 0.04,
                label = paste0("Peak: ", round(time_h, 1), "h")),
            size = 3, hjust = 0, show.legend = FALSE) +

  scale_color_manual(values    = key_colors) +
  scale_linetype_manual(values = c(
    "Bacteria alone"                              = "solid",
    "AB 4 mg/L alone"                             = "dashed",
    "Phage alone"                                 = "dotted",
    "Phage + AB 4 mg/L" = "solid"
  )) +
  scale_x_continuous(breaks = seq(0, 25, 2)) +

  labs(
   # title    = "IMC: Transcriptomics Condition in Context",
  #  subtitle = "Phage + 4 mg/L ATM-AVI (C3) vs each component alone",
    x        = "Time (hours)",
    y        = "Relative heat flow",
    color    = NULL, linetype = NULL
  ) +
  theme_tk +
  theme(
    legend.position  = "top",
    legend.key.width = unit(1.5, "cm"),
    legend.text = element_text(size = 9),
    plot.tag         = element_text(face = "bold", size = 12)
  )
```

```{r}
library(patchwork)
(p_peaks / p_mic_curve) +  plot_annotation(tag_levels =  "A") &
  theme(plot.margin = margin(1, 1, 1, 1),
        axis.title.y.left = element_text(size = 10),
        legend.margin = margin(t = 2, r = 0, b = 2, l = 0, unit = "pt"),
        legend.position = 'top',
        plot.tag = element_text(size=11,face="bold"))

ggsave("./figures/heat_flow.pdf",width=12,height = 10,device = "pdf")
ggsave("./figures/heat_flow.png",width=12,height = 10,device = "png")
```


---

## Derived Metrics from Curves

```{r derived_metrics}
# Calculate TTP, peak value, area under curve, and value at 4h per well
library(DescTools)

derived <- heat %>%
  filter(well != "C2") %>%
  group_by(well, condition, group, ab_conc) %>%
  summarise(
    TTP_h        = time_h[which.max(heat_flow_relative)],
    peak_value   = max(heat_flow_relative),
    value_at_4h  = heat_flow_relative[which.min(abs(time_h - 4))],
    AUC          = AUC(time_h, heat_flow_relative, method = "trapezoid"),
    first_zero   = min(time_h[heat_flow_relative <= 0], default = NA),
    .groups      = "drop"
  ) %>%
  arrange(group, ab_conc)

derived %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(
    col.names = c("Well", "Condition", "Group", "AB (mg/L)",
                  "TTP (h)", "Peak value", "Value at 4h",
                  "AUC", "First crosses zero (h)"),
    caption   = "Derived metrics from digitized heat flow curves"
  )
```

---

## Plot 4: Derived Metrics Comparison

```{r plot_derived, fig.height=5, fig.width=11}
# TTP bar
p_d1 <- derived %>%
  filter(well != "C2") %>%
  mutate(condition = fct_reorder(condition, TTP_h)) %>%
  ggplot(aes(y = condition, x = TTP_h, fill = group)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_vline(xintercept = derived$TTP_h[derived$well == "B1"],
             linetype = "dashed", color = "#2C3E50", alpha = 0.7) +
  scale_fill_manual(values = c(
    "Bacteria"   = "#2C3E50", "Phage alone" = "#2980B9",
    "AB only"    = "#E67E22", "Phage + AB"  = "#8E44AD"
  )) +
  labs(x = "Time to Peak (h)", y = NULL,
       fill = NULL, tag = "D") +
  theme_tk +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 12))

# Heat flow at 4h
p_d2 <- derived %>%
  filter(well != "C2") %>%
  mutate(condition = fct_reorder(condition, -value_at_4h)) %>%
  ggplot(aes(y = condition, x = value_at_4h, fill = group)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_vline(xintercept = 0, color = "grey40",
             linewidth = 0.5, linetype = "dashed") +
  scale_fill_manual(values = c(
    "Bacteria"   = "#2C3E50", "Phage alone" = "#2980B9",
    "AB only"    = "#E67E22", "Phage + AB"  = "#8E44AD"
  )) +
  labs(x = "Relative heat flow at 4h", y = NULL,
       fill = NULL, tag = "E") +
  theme_tk +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 12))

p_d1 | p_d2
```

---

## Manuscript Recommendation

```{r recommendation, echo=FALSE, results='asis'}
ttp_bact   <- derived %>% filter(well == "B1") %>% pull(TTP_h)
ttp_ab4    <- derived %>% filter(well == "B3") %>% pull(TTP_h)
ttp_phage  <- derived %>% filter(well == "C1") %>% pull(TTP_h)
ttp_combo  <- derived %>% filter(well == "C3") %>% pull(TTP_h)
val_bact_4h  <- derived %>% filter(well == "B1") %>% pull(value_at_4h) %>% round(3)
val_combo_4h <- derived %>% filter(well == "C3") %>% pull(value_at_4h) %>% round(3)

cat(glue::glue("
**Key numbers for manuscript:**

- Bacteria alone TTP: {round(ttp_bact,1)}h
- AB 4 mg/L alone TTP: {round(ttp_ab4,1)}h
- Phage alone TTP: {round(ttp_phage,1)}h  
- **Phage + 4 mg/L TTP: {round(ttp_combo,1)}h** (transcriptomics condition)
- Delay vs bacteria alone: +{round(ttp_combo - ttp_bact, 1)}h
- Heat flow at 4h, bacteria: {val_bact_4h}
- Heat flow at 4h, Phage+4mg/L: {val_combo_4h}

**Suggested figure for manuscript:** Plot C (transcriptomics condition in context)
+ Plot B (faceted matched comparison) as a two-panel figure.

**Methods sentence:**
*Heat flow curves were digitized from representative thermograms and used to 
derive time to peak metabolic rate (TTP) and relative heat flow at the 
transcriptomics harvest point (4h). Phage + 4 mg/L ATM-AVI (C3) showed a 
{round(ttp_combo - ttp_bact, 1)}h delay in peak metabolic activity compared 
to bacteria alone, with near-zero heat flow at 4h, supporting selection of 
this timepoint for transcriptomic analysis.*
"))
```
