---
title: "interaction_plots"
author: "Mohammad Razavi"
date: "2026-06-03"
output: html_document
---

```{r libraries}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(scales)
library(RColorBrewer)
library(forcats)
```

---

# 1  Load and prepare data

```{r load-data}
# ── Main transcriptomics table (all time points, all genes) ──────────────────
df <- read.csv("./export/interaction/combo_therapy_responses.csv", stringsAsFactors = FALSE)

# ── 4 h interaction-defined set with functional annotations ─────────────────
int4_long <- read.csv(
  "./export/__response_calss_Interaction-defined_(Non-additive)_4h.csv",
  stringsAsFactors = FALSE
)
```

```{r colour-palette}
# Shared colour scheme used throughout ----------------------------------------
class_colours <- c(
  "Interaction-defined (Non-additive)" = "#C40C0C",
  "Phage-driven (no interaction)"      = "#3396D3",
  "Antibiotic-driven (no interaction)" = "#48A111",
  "Shared (additive, no interaction)"  = "#F4A824",
  "No significant change"              = "#BFC9D1"
)

# Condition colours for lollipop / dot plots
cond_colours <- c(
  phage       = "#3396D3",
  ab          = "#48A111",
  combo       = "#9C27B0",
  interaction = "#C40C0C"
)

# Functional group palette (12 groups — uses Set3 + manual additions)
fg_levels <- c(
  "Anaerobic respiration escape",
  "Proteostasis",
  "Oxidative stress",
  "Outer membrane & porins",
  "Flagellar assembly & motility"
)

set.seed(42)
fg_pal <- c(
  "#D32F2F", "#1565C0", "#2E7D32", "#F57F17",
  "#6A1B9A"
  
)
names(fg_pal) <- fg_levels
```

---

# 3  Panel B — Scatter: interaction coefficient vs antibiotic LFC (4 h)

Genes in the interaction-defined class are highlighted; genes discussed in
the resensitization narrative are labelled.

```{r panel-B-data}
# Work at the 4 h slice
df4 <- df %>% filter(time_point == "4h")

# Genes to label explicitly
label_genes <- c(
  "torA", "torC", "torD",
  "hycE", "hycF",
  "dnaK", "groL", "groS", "hslU", "hslV",
  "ahpC", "ahpF",
  "ompF", "ompC",
  "cadA",
  "znuA",
  "fliA", "tap",
  "acrR", "marA", "acrZ", "ompW"
)

scatter_df <- df4 %>%
  mutate(
    highlight = response_class == "Interaction-defined (Non-additive)",
    label_me  = gene.y %in% label_genes,
    pt_alpha  = case_when(
      highlight ~ 1,
      response_class == "No significant change" ~ 0.15,
      TRUE ~ 0.4
    ),
    pt_size = ifelse(highlight, 2.2, 1.0)
  ) %>%
  filter(!is.na(lfc_interaction), !is.na(lfc_ab))


scatter_df <- scatter_df %>% mutate(response_class=factor(response_class,levels=names(class_colours)))
```

```{r panel-B-plot}
p_scatter <- ggplot(scatter_df,
             aes(x = lfc_ab, y = lfc_interaction,
                 fill = response_class, alpha = pt_alpha, size = pt_size)) +
  # background points
  geom_point(data = filter(scatter_df, !highlight), shape = 21) +
  # interaction-defined points on top
  geom_point(data = filter(scatter_df,  highlight), shape = 21) +
  # reference lines
  geom_hline(yintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  # shaded quadrants (antibiotic-induced & interaction-blunted)
  annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "#FFEBEE", alpha = 0.35) +
  annotate("rect", xmin = -Inf, xmax = 0, ymin = 0, ymax = Inf,
           fill = "#E8F5E9", alpha = 0.35) +
  # quadrant labels
  annotate("text", x = 8, y = -0.5,
           label = "AB-induced\nblunted by phage",
           colour = "#B71C1C", size =3.5, fontface = "bold", hjust = 1) +
  annotate("text", x = -5.5, y = 2.9,
           label = "AB-repressed\nrescued by phage",
           colour = "#1B5E20", size = 3.5, fontface = "bold", hjust = 0) +
  # labels for genes of interest
  geom_label_repel(
    data        = filter(scatter_df, label_me & highlight),
    aes(label   = gene.y),
    size        = 2.8,
    fontface    = "italic",
    fill        = "white",
    colour      = "#B71C1C",
    label.size  = 0.2,
    box.padding = 0.35,
    max.overlaps = 30,
    segment.colour = "grey60",
    segment.size   = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = class_colours) +
  scale_alpha_identity() +
  scale_size_identity() +
  scale_x_continuous(limits = c(-6, 8.5),
                     breaks = seq(-6, 8, by = 2)) +
  scale_y_continuous(limits = c(-3.2, 3.2),
                     breaks = seq(-3, 3, by = 1)) +
  labs(
    #title    = "B  Interaction coefficient vs antibiotic effect (4 h)",
    #subtitle = "Red points: interaction-defined genes  |  shaded: resensitization quadrants",
    x        = expression("Antibiotic LFC (log"[2]*"FC)"),
    y        = expression("Interaction coefficient ("*Delta*"log"[2]*"FC)")
  ) +
  theme_bw(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(colour = "grey40", size = 9),
    axis.text     = element_text(colour = "black"),
    panel.grid.major = element_line(colour = "grey93", linewidth = 0.3),
    legend.position = "top"
  ) + guides(fill=guide_legend(override.aes = list(size=3), nrow=2,title=""))


p_scatter
```


# Panel D — Slope graph: observed vs. sum-of-shrunken-effects for focal genes

Shows the observed combination LFC against the sum of the separately
shrunken phage and antibiotic LFCs, for the most mechanistically important
genes across four functional clusters.

**Note on interpretation**: `lfc_phage` and `lfc_ab` are each apeglm-shrunken
single-coefficient estimates; `lfc_combo` is a separately ashr-shrunken
contrast; `lfc_interaction` is a separately ashr-shrunken interaction
coefficient. Because shrinkage is nonlinear and applied independently to each
quantity, `lfc_phage + lfc_ab` ("Sum*" below) is a *descriptive* reference
point, not an algebraically exact stand-in for the additive-null model — the
visual gap (`lfc_combo - (lfc_phage + lfc_ab)`) can differ appreciably from
the model's actual `lfc_interaction` (e.g. by ~0.6-0.7 log2FC for
torA/torD). The plot itself carries no in-figure footnote (per journal
figure format); this caveat, and the per-gene `lfc_interaction` values in
`pD_df`/`pD_wide` below, belong in the external figure caption/legend text.

```{r panel-D-data}
focal_genes <- tribble(
  ~gene,   ~cluster,

  # ─────────────────────────────────────────────────────────────────────
  # Category 1: Anaerobic respiration escape (tor/hyc)
  # ─────────────────────────────────────────────────────────────────────
  "torA",  "Anaerobic respiration escape",
  "torC",  "Anaerobic respiration escape",
  "torD",  "Anaerobic respiration escape",
  "hycC",  "Anaerobic respiration escape",
  "hycE",  "Anaerobic respiration escape",
  "hycF",  "Anaerobic respiration escape",
  "hycG",  "Anaerobic respiration escape",
  "hycH",  "Anaerobic respiration escape",
  "fdhF",  "Anaerobic respiration escape",
  "dmsD",  "Anaerobic respiration escape",

  # ─────────────────────────────────────────────────────────────────────
  # Category 2: Proteostasis (chaperones + HslVU)
  # ─────────────────────────────────────────────────────────────────────
  "dnaK",  "Proteostasis",
  "dnaJ",  "Proteostasis",
  "groL",  "Proteostasis",
  "groS",  "Proteostasis",
  "htpG",  "Proteostasis",
  "hslU",  "Proteostasis",
  "hslV",  "Proteostasis",
  "ybeZ",  "Proteostasis",

  # ─────────────────────────────────────────────────────────────────────
  # Category 3: Oxidative stress (AhpCF + redox enzymes)
  # ─────────────────────────────────────────────────────────────────────
  "ahpC",  "Oxidative stress",
  "ahpF",  "Oxidative stress",
  "cnoX",  "Oxidative stress",
  "dsbC",  "Oxidative stress",
  "yqhD",  "Oxidative stress",
  "ydjA",  "Oxidative stress",

  # ─────────────────────────────────────────────────────────────────────
  # Category 4: Outer membrane & porins
  # ─────────────────────────────────────────────────────────────────────
  "ompF",  "Outer membrane & porins",
  "ompC",  "Outer membrane & porins",
  "ompW",  "Outer membrane & porins",
  "ompA",  "Outer membrane & porins",
  "ompR",  "Outer membrane & porins",
  "ompX",  "Outer membrane & porins",

)
pD_df <- df4 %>%
  filter(gene.y %in% focal_genes$gene) %>%
  distinct(gene.y, .keep_all = TRUE) %>%
  select(gene = gene.y,
         lfc_phage, lfc_ab, lfc_combo, lfc_interaction) %>%
  # NOTE: sum of two separately-shrunken estimates, NOT an exact additive-null
  # model (shrinkage is nonlinear/independent per coefficient) -- descriptive
  # reference only. See note above; true interaction effect is lfc_interaction.
  mutate(lfc_sum_shrunken = lfc_phage + lfc_ab) %>%
  left_join(focal_genes, by = "gene") %>%
  mutate(
    cluster = factor(cluster, levels = unique(focal_genes$cluster)),
    gene    = factor(gene,    levels = rev(unique(focal_genes$gene)))
  )

# Long for slope: three x positions — Phage, Antibiotic, Sum-of-shrunken, Observed
pD_long <- pD_df %>%
  select(gene, cluster, lfc_phage, lfc_ab, lfc_sum_shrunken, lfc_combo) %>%
  pivot_longer(
    cols      = c(lfc_phage, lfc_ab, lfc_sum_shrunken, lfc_combo),
    names_to  = "condition",
    values_to = "lfc"
  ) %>%
  mutate(
    condition = factor(condition,
                       levels = c("lfc_phage","lfc_ab","lfc_sum_shrunken","lfc_combo"),
                       labels = c("Phage","Antibiotic",
                                  "Sum of shrunken\nestimates (descriptive)","Observed\ncombination")),
    line_type = ifelse(condition %in% c("Sum of shrunken\nestimates (descriptive)","Observed\ncombination"),
                       "focal", "mono")
  )
```

```{r panel-D-plot}
# ── Wide format needed for manual segments ───────────────────────────────────
pD_wide <- pD_df  # already has lfc_phage, lfc_ab, lfc_sum_shrunken, lfc_combo, cluster, gene

p_line <- ggplot() +

  geom_hline(yintercept = 0, colour = "grey55",
             linewidth = 0.4, linetype = "dashed") +

  # ── Segment 1: Phage → Antibiotic (solid) ──────────────────────────────────
  geom_segment(
    data = pD_wide,
    aes(x = 1, xend = 2,
        y = lfc_phage, yend = lfc_ab,
        colour = cluster, group = gene),
    linewidth = 0.6, alpha = 0.75, linetype = "solid"
  ) +

  # ── Segment 2: Antibiotic → Sum of shrunken estimates (dotted, descriptive) ─
  geom_segment(
    data = pD_wide,
    aes(x = 2, xend = 3,
        y = lfc_ab, yend = lfc_sum_shrunken,
        colour = "black", group = gene),
    linewidth = 0.6, alpha = 0.75, linetype = "dashed"
  ) +

  # ── Segment 3: Antibiotic → Observed combination (solid) ───────────────────
  geom_segment(
    data = pD_wide,
    aes(x = 2, xend = 4,
        y = lfc_ab, yend = lfc_combo,
        colour = cluster, group = gene),
    linewidth = 0.6, alpha = 0.75, linetype = "solid"
  ) +

  # ── Visual gap shading (descriptive only -- not equal to lfc_interaction,
  #    see note above) between sum-of-shrunken-estimates and Observed at x=4 ──
  geom_segment(
    data = pD_wide,
    aes(x = 4, xend = 4,
        y = lfc_sum_shrunken, yend = lfc_combo,
        colour = cluster, group = gene),
    linewidth = 2.5, alpha = 0.2
  ) +

  # ── Points ──────────────────────────────────────────────────────────────────
  # Phage
  geom_point(
    data = pD_wide,
    aes(x = 1, y = lfc_phage, fill = cond_colours["phage"]),
    size = 2.5, shape = 21,alpha=0.8
  ) +
  # Antibiotic
  geom_point(
    data = pD_wide,
    aes(x = 2, y = lfc_ab, fill = cond_colours["ab"]),
    size = 2.5, shape = 21,alpha=0.8
  ) +
  # Sum of shrunken estimates (descriptive reference, not the exact additive null)
  geom_point(
    data = pD_wide,
    aes(x = 3, y = lfc_sum_shrunken, colour = "gray40"),
    size = 1, shape = 8, stroke = 1.0
  ) +
  # Observed combination
  geom_point(
    data = pD_wide,
    aes(x = 4, y = lfc_combo, fill = cond_colours["combo"]),
    size = 2.5, shape = 21,alpha=0.8
  ) +

  # ── Gene labels at observed combination. NOTE: the sum-of-shrunken-estimates
  #    point/segment above is a descriptive reference, not the exact
  #    additive-null model -- see figure caption/legend text for the true
  #    (separately-shrunken) interaction coefficient per gene. ─────────────
  geom_text_repel(
    data = pD_wide,
    aes(x = 4, y = lfc_combo,
        label = gene, colour = cluster),
    size           = 4,
    fontface       = "italic",
    nudge_x        = 0.15,
    direction      = "y",
    hjust          = 0,
    segment.size   = 0.3,
    segment.colour = "grey60",
    show.legend    = FALSE
  ) +

  # ── Facet and scales ─────────────────────────────────────────────────────────
  facet_wrap(~ cluster, ncol = 2, scales = "free_y") +

  scale_colour_manual(values =fg_pal , guide = "none") +

  scale_x_continuous(
    breaks = 1:4,
    labels = c("Phage", "Antibiotic", "Sum*", "Combination"),
    limits = c(0.7, 5.0)
  ) +

  labs(
    x = NULL,
    y = expression(log[2]~"fold change")
  ) +

  theme_bw() +
  theme(
    panel.grid.minor   = element_blank(),
    strip.text         = element_text(size = 8.5, face = "bold"),
    axis.text.x        = element_text(size = 8),
    strip.background = element_blank(),
    legend.position = "none"
  )

p_line
```

# 5  Panel D — Four-condition LFC heatmap for focal gene clusters

Shows phage / antibiotic / combo / interaction LFC for the key
resensitization-relevant gene sets discussed in the paper.

```{r panel-heatmap-data}
# Define focal gene sets with their narrative cluster label
focal <- tribble(
  ~gene_name,  ~cluster,

  # ───────────────────────────────────────────────────────────────────────────
  # ANAEROBIC RESCUE (tor/hyc/fdo)
  # ───────────────────────────────────────────────────────────────────────────
  # Original tor operon — TMAO reductase system
  "torA",  "Anaerobic\nrescue\n(tor/hyc)",   # catalytic subunit, strongest interaction signal (Δ -2.83)
  "torC",  "Anaerobic\nrescue\n(tor/hyc)",   # menaquinol electron transport component
  "torD",  "Anaerobic\nrescue\n(tor/hyc)",   # TMAO reductase chaperone, Δ -2.63

  # Original hyc operon — formate hydrogenlyase
  "hycE",  "Anaerobic\nrescue\n(tor/hyc)",   # catalytic large subunit
  "hycF",  "Anaerobic\nrescue\n(tor/hyc)",   # electron transfer, Δ -2.64 (strongest in hyc)
  "hycG",  "Anaerobic\nrescue\n(tor/hyc)",   # Fe-S electron transfer subunit
  "hycH",  "Anaerobic\nrescue\n(tor/hyc)",   # assembly/maturation factor

  # NEW additions — extend the anaerobic escape story
  "hycC",  "Anaerobic\nrescue\n(tor/hyc)",   # membrane-integral subunit of FHL — completes hyc operon picture
  "fdhF",  "Anaerobic\nrescue\n(tor/hyc)",   # formate dehydrogenase H — feeds directly into FHL complex; shows interaction Δ -0.84
  "dmsD",  "Anaerobic\nrescue\n(tor/hyc)",   # DMSO reductase chaperone — third anaerobic electron acceptor pathway
                                              # showing negative interaction; broadens escape route argument beyond tor/hyc alone
  "napG",  "Anaerobic\nrescue\n(tor/hyc)",   # periplasmic nitrate reductase Fe-S component — positive interaction (+0.72)
                                              # contrasts with tor/hyc; shows not all anaerobic genes are blunted equally

  # ───────────────────────────────────────────────────────────────────────────
  # PROTEOSTASIS
  # ───────────────────────────────────────────────────────────────────────────
  "dnaK",  "Proteostasis",   # major Hsp70 chaperone, Δ -1.37
  "dnaJ",  "Proteostasis",   # DnaK co-chaperone, Δ -0.97
  "groL",  "Proteostasis",   # GroEL chaperonin, Δ -0.84
  "groS",  "Proteostasis",   # GroES co-chaperonin, Δ -1.03
  "htpG",  "Proteostasis",   # Hsp90 homologue, Δ -0.74
  "hslU",  "Proteostasis",   # ATPase of HslVU protease, Δ -0.85
  "hslV",  "Proteostasis",   # peptidase of HslVU protease, Δ -0.95

  # NEW additions
  "pspA",  "Proteostasis",   # phage shock protein A — inner membrane stress sentinel
                              # negative interaction (Δ -0.88); links phage membrane stress
                              # to proteostatic response; mechanistically connects phage
                              # lytic activity to the chaperone attenuation story
  "ybeZ",  "Proteostasis",   # uncharacterised but stress-associated, Δ -0.91
                              # co-clusters with chaperones; worth showing as part of the
                              # broader stress response being attenuated
  "rho",   "Proteostasis",   # transcription termination factor Rho — Δ -0.77
                              # Rho-dependent termination shapes stress response mRNA stability;
                              # its attenuation in combination context suggests broader
                              # transcriptional buffering is compromised, not just translation

  # ───────────────────────────────────────────────────────────────────────────
  # OXIDATIVE STRESS
  # ───────────────────────────────────────────────────────────────────────────
  "ahpC",  "Oxidative\nstress",   # alkyl hydroperoxide reductase peroxidase subunit, Δ -0.86
  "ahpF",  "Oxidative\nstress",   # alkyl hydroperoxide reductase reductase subunit, Δ -1.09
  "cnoX",  "Oxidative\nstress",   # chaperedoxin — thioredoxin-like, Δ -0.66
  "dsbC",  "Oxidative\nstress",   # periplasmic disulfide isomerase, Δ -0.54

  # NEW additions
  "yqhD",  "Oxidative\nstress",   # NADPH-dependent aldehyde/lipid peroxide reductase, Δ -0.74
                                   # protects against reactive aldehydes from lipid peroxidation;
                                   # its attenuation links membrane disruption (phage lysis)
                                   # to downstream aldehyde toxicity — extends oxidative story
                                   # beyond H2O2 to lipid oxidation products
  "ydjA",  "Oxidative\nstress",   # DRE2-like iron-sulfur cluster protein, Δ -0.66
                                   # involved in oxidative stress and Fe-S repair;
                                   # connects oxidative stress to iron metabolism disruption


  # ───────────────────────────────────────────────────────────────────────────
  # Flagellar assembly & sigma factor
  # ───────────────────────────────────────────────────────────────────────────
  "fliA",  "Flagella &\nchemotaxis",
  "flgB",  "Flagella &\nchemotaxis",
  "flgD",  "Flagella &\nchemotaxis",
  "flgG",  "Flagella &\nchemotaxis",
  "cheR",  "Flagella &\nchemotaxis",
  "tap",   "Flagella &\nchemotaxis",
  
  # Motile-to-sessile transition
  "rdcB",  "Biofilm\ntransition",
  "mlrA",  "Biofilm\ntransition",
  
  # Fimbriae (negative interaction — contrast)
  "fimB",  "Fimbriae",
  "fimI",  "Fimbriae",
    # ───────────────────────────────────────────────────────────────────────────
  # OUTER MEMBRANE & PORINS
  # ───────────────────────────────────────────────────────────────────────────
  "ompF",  "Outer\nmembrane",   # frameshifted — positive interaction (+1.26) is regulatory readout only
  "ompC",  "Outer\nmembrane",   # sole functional general diffusion porin; phage-driven repression (lfc_phage -1.74)
  "ompW",  "Outer\nmembrane",   # massive phage-driven repression (lfc_phage -5.63) — likely receptor downregulation
  "ompA",  "Outer\nmembrane",   # structural OMP; shared additive induction under both conditions

  # NEW additions
  "ompR",  "Outer\nmembrane",   # EnvZ/OmpR two-component response regulator — master regulator of ompC/ompF
                                 # shared additive induction (lfc_phage +0.96, lfc_ab +0.45)
                                 # showing both agents activate the OmpR regulon independently;
                                 # yet ompC goes down — demonstrating post-OmpR override by phage
                                 # This paradox is worth showing directly in the figure
  "ompX",  "Outer\nmembrane",   # antibiotic-driven only (lfc_ab +0.67, no phage effect, no interaction)
                                 # small OMP associated with reduced permeability;
                                 # its antibiotic-specific induction with no phage effect
                                 # contrasts with ompC and ompW — shows selectivity of phage membrane effects

  # ───────────────────────────────────────────────────────────────────────────
  # ACID STRESS & AMINO ACID STRESS
  # ───────────────────────────────────────────────────────────────────────────
  #"cadA",  "Acid\nstress",   # lysine decarboxylase — primary acid resistance, Δ -1.21
  #"lysU",  "Acid\nstress",   # stress-inducible lysyl-tRNA synthetase, Δ -0.73

  # NEW additions
  #"asnB",  "Acid\nstress",   # asparagine synthetase B — amino acid stress response, Δ -0.73
                              # induced by aztreonam-avibactam alone; blunted in combination;
                              # reflects impaired nitrogen stress buffering
  #"cysK",  "Acid\nstress",   # cysteine synthase A — Δ -0.67
                              # induced under cell wall/envelope stress and sulfur limitation;
                              # connects amino acid metabolism to envelope damage response

  # ───────────────────────────────────────────────────────────────────────────
  # ZINC HOMEOSTASIS (NDM-5 cofactor axis)
  # ───────────────────────────────────────────────────────────────────────────
  #"znuA",  "Zinc\nhomeostasis",   # high-affinity zinc ABC transporter periplasmic binding protein, Δ +1.50
  #"znuC",  "Zinc\nhomeostasis",   # zinc ABC transporter ATPase, Δ +0.59

  # NEW addition
  #"feoA",  "Zinc\nhomeostasis",   # ferrous iron uptake system component, Δ -0.80
                                   # iron and zinc homeostasis are co-regulated under oxidative stress;
                                   # feoA negative interaction alongside znuA positive interaction
                                   # shows the metal homeostasis axis is broadly dysregulated —
                                   # not just zinc but iron import is disrupted, relevant to
                                   # Fe-S cluster repair and NDM-5 active site integrity
  
  # ───────────────────────────────────────────────────────────────────────────
  # MAR/ACR REGULON (no interaction — contrast group)
  # ───────────────────────────────────────────────────────────────────────────
  "marA",  "Mar/Acr\nregulon",   # phage-driven master activator (lfc_phage +1.98); no interaction
  "marR",  "Mar/Acr\nregulon",   # phage-driven autorepressor (lfc_phage +1.68); no interaction
  "acrR",  "Mar/Acr\nregulon",   # phage-driven repression of acrAB repressor (lfc_phage -1.31); no interaction
  "acrA",  "Mar/Acr\nregulon",   # phage-driven (lfc_phage +0.61); no interaction
  "acrB",  "Mar/Acr\nregulon",   # no significant change under any condition — key contrast point
  "acrZ",  "Mar/Acr\nregulon",   # strongly phage-repressed (lfc_phage -2.61); no interaction
  "tolC",  "Mar/Acr\nregulon",   # phage-driven (lfc_phage +0.87); no interaction

  # NEW additions to mar/acr group
  "acrD",  "Mar/Acr\nregulon",   # aminoglycoside-specific RND efflux; phage-repressed (lfc_phage -1.20)
                                   # irrelevant to aztreonam mechanism but confirms phage broadly
                                   # remodels efflux landscape independently of antibiotic
  "acrE",  "Mar/Acr\nregulon",   # secondary efflux AcrEF — strongly phage-induced (lfc_phage +2.51)
                                   # diverges from acrF (phage-repressed) on same operon;
                                   # shows dissociated regulation of AcrEF components;
                                   # no interaction for either — efflux not part of synergy
  "marB",  "Mar/Acr\nregulon",   # no significant change — completes the mar locus picture;
                                   # together with marA/marR shows the full operon is phage-driven
                                   # with zero antibiotic contribution, confirming YRIK renders
                                   # the bacteria transcriptionally blind to aztreonam-avibactam
                                   # at the mar sensing level
  "marC",   "Mar/Acr\nregulon"    # inner membrane protein, phage-repressed (lfc_phage -0.72);
                                   # no interaction; completes the operon representation
)
heat_df <- df4 %>%
  filter(gene.y %in% focal$gene_name) %>%
  select(gene_name = gene.y,
         Phage       = lfc_phage,
         Antibiotic  = lfc_ab,
         Combination = lfc_combo,
         Interaction = lfc_interaction) %>%
  distinct(gene_name, .keep_all = TRUE) %>%
  left_join(focal, by = "gene_name") %>%
  pivot_longer(cols = c(Phage, Antibiotic, Combination, Interaction),
               names_to = "condition", values_to = "lfc") %>%
  mutate(
    condition = factor(condition,
                       levels = c("Phage","Antibiotic","Combination","Interaction")),
    cluster   = factor(cluster, levels = unique(focal$cluster)),
    gene_name = factor(gene_name,
                       levels = rev(unique(focal$gene_name)))
  )
```

```{r panel-heatmap-plot}
lfc_limit <- 8

p_heat <- ggplot(heat_df,
             aes(x = condition, y = gene_name, fill = lfc)) +
  geom_tile(colour = "gray70", linewidth = 0.4) +
  geom_text(aes(label = round(lfc, 1)),
            size = 3, colour = "black",
            fontface = "bold",
            data = filter(heat_df, abs(lfc) >= 0.8)) +
  facet_grid(cluster ~ ., scales = "free_y",
             space = "free_y", switch = "y",) +
  scale_fill_gradientn(
    colours = c("#1565C0", "#42A5F5", "#E3F2FD",
                "white",
                "#FFEBEE", "#EF5350", "#B71C1C"),
    values  = rescale(c(-lfc_limit, -2, -0.5, 0, 0.5, 2, lfc_limit)),
    limits  = c(-lfc_limit, lfc_limit),
    oob     = squish,
    name    = expression(log[2]*"FC"),
    guide   = guide_colourbar(
      direction      = "horizontal",
      title.position = "left",
      title.vjust    = 1,
      barwidth       = unit(6, "cm"),
      barheight      = unit(0.4, "cm")
    )
  ) +
  # vertical line separating mono-therapies from interaction column
  geom_vline(xintercept = 3.5, colour = "grey30",
             linewidth = 0.8, linetype = "solid") +
  labs(
    #title    = "D  LFC heatmap — focal resensitization gene clusters (4 h)",
    #subtitle = "Left of line: observed LFC per condition; right: interaction coefficient (Δlog₂FC)",
    x        = NULL,
    y        = NULL
  ) +
  theme_classic(base_size = 9.5) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    plot.subtitle      = element_text(colour = "grey40", size = 8),
    axis.text.y        = element_text(size = 8, face = "italic"),
    axis.text.x        = element_text(size = 9, colour = "black",
                                      face = c("plain","plain","plain","bold")),
    strip.text.y.left  = element_text(angle = 0, hjust = 0.5,
                                      size = 9, face = "bold"),
    strip.placement    = "outside",
    strip.background   = element_rect(fill = "#FAFAFA", colour = "gray70"),
    panel.grid         = element_blank(),
    panel.spacing      = unit(0.15, "lines"),
    legend.position = "top",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.justification = "center",
  )

p_heat
```

---

```{r}


((p_scatter / p_line) | p_heat )+ plot_layout(widths = c(1.2,0.8)) + 
  plot_annotation(tag_levels = list(c("A","C","B")))

ggsave("./figures/interaction_panel.pdf",width=12,height = 13,device="pdf")

# JPEG copy (300 dpi, DPI metadata embedded) replaces the PNG for journal submission
magick::image_write(
  magick::image_read_pdf("./figures/interaction_panel.pdf", density = 300),
  path = "./figures/interaction_panel.jpg", format = "jpg", quality = 95, density = "300x300"
)
```

