---
title: "Phage-Antibiotic Combination: Interaction-Defined Transcriptional Responses"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = T,
  message = T

)
```

```{r libraries}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(scales)
library(RColorBrewer)
library(forcats)
```

---

# 1  Load and prepare data

```{r load-data}
# ── Main transcriptomics table (all time points, all genes) ──────────────────
df <- read.csv("./export/interaction/combo_therapy_responses.csv", stringsAsFactors = FALSE)

# ── 4 h interaction-defined set with functional annotations ─────────────────
int4_long <- read.csv(
  "./export/__response_calss_Interaction-defined_(Non-additive)_4h.csv",
  stringsAsFactors = FALSE
)
```

```{r colour-palette}
# Shared colour scheme used throughout ----------------------------------------
class_colours <- c(
  "No significant change"              = "#BDBDBD",
  "Phage-driven (no interaction)"      = "#4CAF50",
  "Antibiotic-driven (no interaction)" = "#FF9800",
  "Shared (additive, no interaction)"  = "#2196F3",
  "Interaction-defined (Non-additive)" = "#E53935"
)

# Condition colours for lollipop / dot plots
cond_colours <- c(
  phage       = "#4CAF50",
  ab          = "#FF9800",
  combo       = "#9C27B0",
  interaction = "#E53935"
)

# Functional group palette (12 groups — uses Set3 + manual additions)
fg_levels <- c(
  "Anaerobic respiration & redox metabolism",
  "Protein quality control & stress",
  "Oxidative stress response",
  "Outer membrane permeability & porins",
  "Flagellar assembly & motility",
  "Acid stress & amino acid metabolism",
  "Polyamine metabolism",
  "Zinc acquisition & metal homeostasis",
  "DNA replication & repair",
  "Biofilm & surface regulation",
  "Adhesion & fimbriae",
  "Efflux & antibiotic resistance",
  "Other"
)

set.seed(42)
fg_pal <- c(
  "#D32F2F", "#1565C0", "#2E7D32", "#F57F17",
  "#6A1B9A", "#AD1457", "#00695C", "#E65100",
  "#283593", "#4E342E", "#37474F", "#558B2F",
  "#78909C"
)
names(fg_pal) <- fg_levels
```

---

# 2  Panel A — Stacked bar: response class composition across time points

```{r panel-A-data}
# Count genes per response class × time point
bar_df <- df %>%
  count(time_point, response_class) %>%
  mutate(
    time_point    = factor(time_point, levels = c("1h", "4h", "24h")),
    response_class = factor(response_class, levels = names(class_colours))
  )
```

```{r panel-A-plot}
pA <- ggplot(bar_df, aes(x = time_point, y = n, fill = response_class)) +
  geom_bar(stat = "identity", width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = ifelse(n > 50, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = class_colours,
    name   = "Response class",
    guide  = guide_legend(ncol = 1, keywidth = 0.8, keyheight = 0.8)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.04))) +
  labs(
    title    = "A  Response class composition",
    subtitle = "All genes across three time points",
    x        = "Time point",
    y        = "Number of genes"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    plot.subtitle    = element_text(colour = "grey40", size = 9),
    legend.position  = "right",
    legend.title     = element_text(face = "bold", size = 9),
    legend.text      = element_text(size = 8),
    axis.text        = element_text(colour = "black"),
    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.3)
  )

pA
```

---

# 3  Panel B — Scatter: interaction coefficient vs antibiotic LFC (4 h)

Genes in the interaction-defined class are highlighted; genes discussed in
the resensitization narrative are labelled.

```{r panel-B-data}
# Work at the 4 h slice
df4 <- df %>% filter(time_point == "4h")

# Genes to label explicitly
label_genes <- c(
  "torA", "torC", "torD",
  "hycE", "hycF",
  "dnaK", "groL", "groS", "hslU", "hslV",
  "ahpC", "ahpF",
  "ompF", "ompC",
  "cadA",
  "znuA",
  "fliA", "tap",
  "acrR", "marA", "acrZ", "ompW"
)

scatter_df <- df4 %>%
  mutate(
    highlight = response_class == "Interaction-defined (Non-additive)",
    label_me  = gene.y %in% label_genes,
    pt_alpha  = case_when(
      highlight ~ 1,
      response_class == "No significant change" ~ 0.15,
      TRUE ~ 0.4
    ),
    pt_size = ifelse(highlight, 2.2, 1.0)
  ) %>%
  filter(!is.na(lfc_interaction), !is.na(lfc_ab))
```

```{r panel-B-plot, fig.height=7}
pB <- ggplot(scatter_df,
             aes(x = lfc_ab, y = lfc_interaction,
                 colour = response_class, alpha = pt_alpha, size = pt_size)) +
  # background points
  geom_point(data = filter(scatter_df, !highlight), shape = 16) +
  # interaction-defined points on top
  geom_point(data = filter(scatter_df,  highlight), shape = 16) +
  # reference lines
  geom_hline(yintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  # shaded quadrants (antibiotic-induced & interaction-blunted)
  annotate("rect", xmin = 0, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "#FFEBEE", alpha = 0.35) +
  annotate("rect", xmin = -Inf, xmax = 0, ymin = 0, ymax = Inf,
           fill = "#E8F5E9", alpha = 0.35) +
  # quadrant labels
  annotate("text", x = 7.5, y = -3.1,
           label = "AB-induced\nblunted by phage",
           colour = "#B71C1C", size = 2.8, fontface = "italic", hjust = 1) +
  annotate("text", x = -3.0, y = 2.9,
           label = "AB-repressed\nrescued by phage",
           colour = "#1B5E20", size = 2.8, fontface = "italic", hjust = 0) +
  # labels for genes of interest
  geom_label_repel(
    data        = filter(scatter_df, label_me & highlight),
    aes(label   = gene.y),
    size        = 2.8,
    fontface    = "italic",
    fill        = "white",
    colour      = "#B71C1C",
    label.size  = 0.2,
    box.padding = 0.35,
    max.overlaps = 30,
    segment.colour = "grey60",
    segment.size   = 0.3,
    show.legend = FALSE
  ) +
  scale_colour_manual(values = class_colours, guide = "none") +
  scale_alpha_identity() +
  scale_size_identity() +
  scale_x_continuous(limits = c(-6, 8.5),
                     breaks = seq(-6, 8, by = 2)) +
  scale_y_continuous(limits = c(-3.2, 3.2),
                     breaks = seq(-3, 3, by = 1)) +
  labs(
    title    = "B  Interaction coefficient vs antibiotic effect (4 h)",
    subtitle = "Red points: interaction-defined genes  |  shaded: resensitization quadrants",
    x        = expression("Antibiotic LFC (log"[2]*"FC)"),
    y        = expression("Interaction coefficient (\u0394log"[2]*"FC)")
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(colour = "grey40", size = 9),
    axis.text     = element_text(colour = "black"),
    panel.grid.major = element_line(colour = "grey93", linewidth = 0.3)
  )

pB
```


---

# 4  Panel C — Dot plot: interaction coefficients by functional group

All 113 interaction-defined genes at 4 h, coloured by functional group,
sorted by interaction coefficient within group.

```{r panel-C-data}
# Extract interaction rows and join functional group
inter_genes <- int4_long %>%
  filter(group2 == "I") %>%          # interaction coefficient rows only
  select(gene_name, functional_group, lfc_interaction = LFC) %>%
  mutate(
    functional_group = case_when(
      is.na(functional_group) ~ "Other",
      functional_group %in% fg_levels ~ functional_group,
      TRUE ~ "Other"
    ),
    functional_group = factor(functional_group, levels = fg_levels),
    direction = ifelse(lfc_interaction < 0, "Negative\n(blunted)", "Positive\n(rescued)")
  ) %>%
  # remove duplicates (some genes appear twice due to Prokka annotation)
  distinct(gene_name, .keep_all = TRUE) %>%
  arrange(functional_group, lfc_interaction) %>%
  mutate(gene_name = fct_inorder(gene_name))
```

```{r panel-C-plot}
pC <- ggplot(inter_genes,
             aes(x = lfc_interaction, y = gene_name,
                 colour = functional_group)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_segment(aes(x = 0, xend = lfc_interaction,
                   y = gene_name, yend = gene_name),
               linewidth = 0.5, alpha = 0.6) +
  geom_point(size = 2.8) +
  facet_grid(functional_group ~ .,
             scales = "free_y", space = "free_y",
             switch = "y") +
  scale_colour_manual(values = fg_pal, guide = "none") +
  scale_x_continuous(breaks = seq(-3, 3, by = 1),
                     limits = c(-3.2, 3.2)) +
  labs(
    title    = "C  Interaction coefficients by functional group (4 h)",
    subtitle = "Each point = one gene; left = phage blunts antibiotic induction; right = phage rescues antibiotic repression",
    x        = expression("Interaction coefficient (\u0394log"[2]*"FC)"),
    y        = NULL
  ) +
  theme_bw(base_size = 9.5) +
  theme(
    plot.title        = element_text(face = "bold", size = 12),
    plot.subtitle     = element_text(colour = "grey40", size = 8),
    strip.text.y.left = element_text(angle = 0, hjust = 1,
                                     size = 7.5, face = "bold"),
    strip.placement   = "outside",
    strip.background  = element_rect(fill = "grey95", colour = NA),
    panel.grid.major.x = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 7.5, face = "italic"),
    axis.text.x        = element_text(colour = "black"),
    panel.spacing      = unit(0.15, "lines")
  )

pC
```

---

# 5  Panel D — Four-condition LFC heatmap for focal gene clusters

Shows phage / antibiotic / combo / interaction LFC for the key
resensitization-relevant gene sets discussed in the paper.

```{r panel-D-data}
# Define focal gene sets with their narrative cluster label
focal <- tribble(
  ~gene_name,  ~cluster,
  # Anaerobic escape
  "torA",  "Anaerobic\nrescue\n(tor/hyc)",
  "torC",  "Anaerobic\nrescue\n(tor/hyc)",
  "torD",  "Anaerobic\nrescue\n(tor/hyc)",
  "hycE",  "Anaerobic\nrescue\n(tor/hyc)",
  "hycF",  "Anaerobic\nrescue\n(tor/hyc)",
  "hycG",  "Anaerobic\nrescue\n(tor/hyc)",
  "hycH",  "Anaerobic\nrescue\n(tor/hyc)",
  # Proteostasis
  "dnaK",  "Proteostasis",
  "dnaJ",  "Proteostasis",
  "groL",  "Proteostasis",
  "groS",  "Proteostasis",
  "htpG",  "Proteostasis",
  "hslU",  "Proteostasis",
  "hslV",  "Proteostasis",
  # Oxidative stress
  "ahpC",  "Oxidative\nstress",
  "ahpF",  "Oxidative\nstress",
  "cnoX",  "Oxidative\nstress",
  "dsbC",  "Oxidative\nstress",
  # Outer membrane / porins
  "ompF",  "Outer\nmembrane",
  "ompC",  "Outer\nmembrane",
  "ompW",  "Outer\nmembrane",
  "ompA",  "Outer\nmembrane",
  # Acid stress
  "cadA",  "Acid\nstress",
  "lysU",  "Acid\nstress",
  # Zinc
  "znuA",  "Zinc\nhomeostasis",
  "znuC",  "Zinc\nhomeostasis",
  # Efflux / mar
  "marA",  "Mar/Acr\nregulon",
  "marR",  "Mar/Acr\nregulon",
  "acrR",  "Mar/Acr\nregulon",
  "acrA",  "Mar/Acr\nregulon",
  "acrB",  "Mar/Acr\nregulon",
  "acrZ",  "Mar/Acr\nregulon",
  "tolC",  "Mar/Acr\nregulon"
)

heat_df <- df4 %>%
  filter(gene.y %in% focal$gene_name) %>%
  select(gene_name = gene.y,
         Phage       = lfc_phage,
         Antibiotic  = lfc_ab,
         Combination = lfc_combo,
         Interaction = lfc_interaction) %>%
  distinct(gene_name, .keep_all = TRUE) %>%
  left_join(focal, by = "gene_name") %>%
  pivot_longer(cols = c(Phage, Antibiotic, Combination, Interaction),
               names_to = "condition", values_to = "lfc") %>%
  mutate(
    condition = factor(condition,
                       levels = c("Phage","Antibiotic","Combination","Interaction")),
    cluster   = factor(cluster, levels = unique(focal$cluster)),
    gene_name = factor(gene_name,
                       levels = rev(unique(focal$gene_name)))
  )
```

```{r panel-D-plot, fig.height=9, fig.width=10}
lfc_limit <- 8

pD <- ggplot(heat_df,
             aes(x = condition, y = gene_name, fill = lfc)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = round(lfc, 1)),
            size = 2.5, colour = "white",
            fontface = "bold",
            data = filter(heat_df, abs(lfc) >= 0.8)) +
  facet_grid(cluster ~ ., scales = "free_y",
             space = "free_y", switch = "y") +
  scale_fill_gradientn(
    colours  = c("#1565C0", "#42A5F5", "#E3F2FD",
                 "white",
                 "#FFEBEE", "#EF5350", "#B71C1C"),
    values   = rescale(c(-lfc_limit, -2, -0.5, 0, 0.5, 2, lfc_limit)),
    limits   = c(-lfc_limit, lfc_limit),
    oob      = squish,
    name     = expression(log[2]*"FC"),
    guide    = guide_colourbar(
      barwidth  = 0.6,
      barheight = 8,
      title.position = "top",
      title.hjust    = 0.5
    )
  ) +
  # vertical line separating mono-therapies from interaction column
  geom_vline(xintercept = 3.5, colour = "grey30",
             linewidth = 0.8, linetype = "solid") +
  labs(
    title    = "D  LFC heatmap — focal resensitization gene clusters (4 h)",
    subtitle = "Left of line: observed LFC per condition; right: interaction coefficient (Δlog₂FC)",
    x        = NULL,
    y        = NULL
  ) +
  theme_bw(base_size = 9.5) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    plot.subtitle      = element_text(colour = "grey40", size = 8),
    axis.text.y        = element_text(size = 8, face = "italic"),
    axis.text.x        = element_text(size = 9, colour = "black",
                                      face = c("plain","plain","plain","bold")),
    strip.text.y.left  = element_text(angle = 0, hjust = 1,
                                      size = 7.5, face = "bold"),
    strip.placement    = "outside",
    strip.background   = element_rect(fill = "grey95", colour = NA),
    panel.grid         = element_blank(),
    panel.spacing      = unit(0.15, "lines"),
    legend.position    = "right"
  )

pD
```

---

# 6  Panel E — Selected gene lollipop: four-condition profile

Focuses on the strongest interaction signals, showing the four LFC values
side-by-side for intuitive interpretation of the non-additivity.

```{r panel-E-data}
top_genes <- c("torA","torC","torD","hycF","hycE",
               "dnaK","groL","hslU","ahpF","ahpC",
               "cadA","ompF","ompC","znuA","marA","acrZ")

lollipop_df <- df4 %>%
  filter(gene.y %in% top_genes) %>%
  select(gene_name = gene.y,
         phage       = lfc_phage,
         ab          = lfc_ab,
         combo       = lfc_combo,
         interaction = lfc_interaction) %>%
  distinct(gene_name, .keep_all = TRUE) %>%
  pivot_longer(-gene_name, names_to = "condition", values_to = "lfc") %>%
  mutate(
    condition = factor(condition,
                       levels = c("phage","ab","combo","interaction"),
                       labels = c("Phage","Antibiotic","Combination","Interaction \u0394")),
    gene_name = factor(gene_name, levels = rev(top_genes))
  )
```

```{r panel-E-plot, fig.height=7}
pE <- ggplot(lollipop_df,
             aes(x = lfc, y = gene_name, colour = condition)) +
  geom_vline(xintercept = 0, colour = "grey40",
             linewidth = 0.5, linetype = "dashed") +
  geom_segment(aes(x = 0, xend = lfc,
                   y = gene_name, yend = gene_name),
               linewidth = 0.8, alpha = 0.7) +
  geom_point(size = 3) +
  facet_wrap(~ condition, nrow = 1) +
  scale_colour_manual(values = cond_colours, guide = "none") +
  scale_x_continuous(breaks = seq(-4, 8, by = 2)) +
  labs(
    title    = "E  Four-condition LFC profiles — top interaction genes (4 h)",
    subtitle = "Each row = one gene; panels show observed LFC per condition and the interaction coefficient",
    x        = expression(log[2]*"FC"),
    y        = NULL
  ) +
  theme_bw(base_size = 10) +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    plot.subtitle    = element_text(colour = "grey40", size = 9),
    strip.text       = element_text(face = "bold", size = 10),
    strip.background = element_rect(fill = "grey95", colour = NA),
    axis.text.y      = element_text(size = 9, face = "italic"),
    axis.text.x      = element_text(colour = "black"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey93")
  )

pE
```

---

# 7  Compose multi-panel figure

## Figure 1: Overview panels (A + B)

```{r fig1, fig.width=14, fig.height=6}
fig1 <- pA + pB +
  plot_annotation(
    title   = "Figure 1 | Response class landscape and interaction overview",
    theme   = theme(
      plot.title = element_text(face = "bold", size = 13, hjust = 0)
    )
  ) +
  plot_layout(widths = c(1, 1.6))

fig1
```

```{r save-fig1}
ggsave("Figure1_overview.pdf",  fig1, width = 14, height = 6,  device = "pdf")
ggsave("Figure1_overview.png",  fig1, width = 14, height = 6,  dpi = 300)
```

## Figure 2: Interaction-defined gene details (C + D)

```{r fig2, fig.width=18, fig.height=11}
fig2 <- pC + pD +
  plot_annotation(
    title   = "Figure 2 | Interaction-defined gene clusters and LFC heatmap (4 h)",
    theme   = theme(
      plot.title = element_text(face = "bold", size = 13, hjust = 0)
    )
  ) +
  plot_layout(widths = c(1.1, 1))

fig2
```

```{r save-fig2}
ggsave("Figure2_interaction_clusters.pdf", fig2, width = 18, height = 11, device = "pdf")
ggsave("Figure2_interaction_clusters.png", fig2, width = 18, height = 11, dpi = 300)
```

## Figure 3: Four-condition lollipop (E — standalone)

```{r fig3, fig.width=13, fig.height=7}
fig3 <- pE +
  plot_annotation(
    title = "Figure 3 | Four-condition LFC profiles — top interaction genes (4 h)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 13, hjust = 0)
    )
  )

fig3
```

```{r save-fig3}
ggsave("Figure3_lollipop.pdf", fig3, width = 13, height = 7, device = "pdf")
ggsave("Figure3_lollipop.png", fig3, width = 13, height = 7, dpi = 300)
```

## Combined 5-panel figure (publication layout)

```{r fig-combined, fig.width=18, fig.height=20}
fig_combined <-
  (pA | pB) /
  (pC | pD) /
  pE +
  plot_annotation(
    title   = "Phage–antibiotic combination: interaction-defined transcriptional responses",
    caption = "LFC: log₂ fold change relative to untreated control. Interaction coefficient (Δlog₂FC): departure of combination response from additive expectation.",
    tag_levels = "A",
    theme = theme(
      plot.title   = element_text(face = "bold", size = 14, hjust = 0),
      plot.caption = element_text(colour = "grey50", size = 8, hjust = 0)
    )
  ) +
  plot_layout(heights = c(1, 1.6, 0.9))

fig_combined
```

```{r save-combined}
ggsave("Figure_combined.pdf", fig_combined,
       width = 18, height = 20, device = "pdf")
ggsave("Figure_combined.png", fig_combined,
       width = 18, height = 20, dpi = 300)
```

---

# 8  Session info

```{r session-info}
sessionInfo()
```
