---
title: "Time-Kill Curve Analysis: Phage-Antibiotic Synergy"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 150)
```

## Libraries & Parameters

```{r libraries}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(scales)
library(ggrepel)
library(patchwork)
library(rstatix)
library(ggpubr)
# ============================================================
# USER PARAMETERS
# ============================================================
data_dir <- "../../time_kill_data/"
# Input file
input_file <- file.path(data_dir,"time_kill.csv")

# Initial inoculum at t=0 (CFU/mL)
initial_inoculum <- 1e6

# Detection limit (for values of 0, replace with this)
detection_limit <- 100

# Log10 killing threshold for "synergy" or "bactericidal"
bactericidal_threshold <- 4  # ≥3 log10 reduction = bactericidal

# Colors per condition group
# color_palette <- c(
#   "BC-CTR"       = "#2C3E50",   # Dark grey - bacteria control
#   "BC-8mg/L"     = "#E74C3C",   # Red       - antibiotic 8 mg/L
#   "BC-4mg/L"     = "#E67E22",   # Orange    - antibiotic 4 mg/L
#   "BC-2mg/L"     = "#F39C12",   # Yellow    - antibiotic 2 mg/L
#   "BC-1mg/L"     = "#F1C40F",   # Light yellow
#   "BC-0.5mg/L"   = "#FAD7A0",   # Very light
#   "PH-CTRL"      = "#2980B9",   # Blue      - phage alone
#   "PHBC-4mg/L"   = "#8E44AD",   # Purple    - phage + 4 mg/L
#   "PHBC-2mg/L"   = "#9B59B6",   # Medium purple
#   "PHBC-1mg/L"   = "#BB8FCE",   # Light purple
#   "PHBC-0.5mg/L" = "#D7BDE2"    # Very light purple
# )

# color_palette <- c(
#   "BC-CTR"       = "#2C3E50",   # Dark grey - bacteria control
#   "BC-8mg/L"     = "#E74C3C",   # Red       - antibiotic 8 mg/L
#   "BC-4mg/L"     = "#E67E22",   # Orange    - antibiotic 4 mg/L
#   "BC-2mg/L"     = "#F39C12",   # Yellow    - antibiotic 2 mg/L
#   "BC-1mg/L"     = "#F1C40F",   # Light yellow
#   "BC-0.5mg/L"   = "#FAD7A0",   # Very light
#   "PH-CTRL"      = "#2980B9",   # Blue      - phage alone
#   "PHBC-4mg/L"   = "#8E44AD",   # Purple    - phage + 4 mg/L
#   "PHBC-2mg/L"   = "#9B59B6",   # Medium purple
#   "PHBC-1mg/L"   = "#BB8FCE",   # Light purple
#   "PHBC-0.5mg/L" = "#D7BDE2"    # Very light purple
# )
color_palette <- c(
  "BC-CTR"       = "#2C3E50",   
  "BC-8mg/L"     = "#E74C3C",  
  "BC-4mg/L"     = "#8A5F41",  
  "BC-2mg/L"     = "#F39C12",  
  "BC-1mg/L"     = "#9FA1FF",   
  "BC-0.5mg/L"   = "#6FCF97",   
  "PH-CTRL"      = "#2C3E50",   
  "PHBC-4mg/L"   = "#8A5F41",   
  "PHBC-2mg/L"   = "#F39C12",  
  "PHBC-1mg/L"   = "#9FA1FF",   
  "PHBC-0.5mg/L" = "#6FCF97"  
)

shape_colors <- c(   
  "BC-8mg/L"     = "#E74C3C",  
  "BC-4mg/L"     = "#8A5F41",  
  "BC-2mg/L"     = "#F39C12",  
  "BC-1mg/L"     = "#9FA1FF",   
  "BC-0.5mg/L"   = "#6FCF97",
  "BC-CTR"       = "#2C3E50") 

# Shared theme
theme_tk <- theme_bw(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(r=20,l=10),
        plot.caption = element_text(hjust = 0,size=12),
              axis.text.x=element_text(size=12,hjust=0.5,vjust = 0.5),
              axis.text.y=element_text(size=12),
              axis.title.y = element_text(size=14),
              axis.title.x = element_text(size=13),
              strip.background = element_blank(), strip.placement = "outside",strip.text=element_text(size=12),
              panel.grid.major.x = element_line(linewidth =0.3,linetype = "solid"),
              panel.grid.major.y = element_line(linetype = "dashed", colour = "gray80"),
              panel.grid.minor.y = element_line(linetype = "dotted", colour = "gray80"))

```

## Data Import & Preprocessing

```{r data_prep}
# Read data
raw <- read.csv(input_file, stringsAsFactors = FALSE, check.names = FALSE)

# Add t=0 (initial inoculum for each row)
raw <- raw %>% mutate(`0h` = initial_inoculum)

# Convert to long format
long <- raw %>%
  pivot_longer(
    cols = matches("^[0-9]+h$"),
    names_to = "timepoint",
    values_to = "cfu"
  ) %>%
  mutate(
    time = as.numeric(str_remove(timepoint, "h")),
    # Replace 0 and NA with detection limit
    cfu = ifelse( cfu == 0, detection_limit, cfu),
    log_cfu = log10(cfu),
    # Group variables
    phage = factor(phage, levels = c(0, 1), labels = c("No Phage", "Phage")),
    antibiotic = factor(antibiotic)
  )

# Summarise: mean ± SD per condition
summary_df <- long %>%
  group_by(name, phage, antibiotic, time) %>%
  summarise(
    mean_log = mean(log_cfu),
    sd_log   = sd(log_cfu),
    se_log   = sd_log / sqrt(n()),
    n        = n(),
    .groups  = "drop"
  ) %>%
  # Log10 reduction from initial inoculum
  mutate(log_reduction = log10(initial_inoculum) - mean_log)

# Log reduction at 24h only
final_df <- summary_df %>%
  filter(time == 24) %>%
  arrange(log_reduction)

cat("Conditions:", unique(summary_df$name), "\n")
cat("Time points:", sort(unique(summary_df$time)), "\n")
cat("Replicates per group: 3\n")
```
---

## Plot 2: Time-Kill Curves (Faceted: Phage vs No Phage)

> Compares antibiotic-only vs phage+antibiotic conditions side by side.

```{r plot2_facet, fig.height=6, fig.width=12}

  
 library(patchwork)

# ============================================================
# FUNCTION: draw one panel by phage group
# ============================================================

plot_timekill <- function(data, 
                          phage_group,    # "No Phage" or "Phage"
                          panel_label,    # e.g. "(A) Antibiotic Only"
                          show_legend  = FALSE,
                          show_x_title = FALSE) {
  
  df <- data %>% filter(phage == phage_group)
  
  p <- ggplot(df, aes(x = time, y = mean_log,
                      color = name, fill = name,
                      group = name, shape = antibiotic)) +
    
    geom_ribbon(aes(ymin = mean_log - sd_log,
                    ymax = mean_log + sd_log),
                alpha = 0.15, color = NA) +
    geom_line(linewidth = 0.3) +
    geom_point(size = 2.5, alpha = 0.6) +
    
    geom_hline(yintercept = log10(detection_limit),
               linetype = "dotted", color = "red") +
    annotate("text", x = 0.05, y = log10(detection_limit) + 0.15,
             label = "Detection limit", size = 3,
             color = "red", hjust = 0) +
    
    scale_color_manual(values = color_palette) +
    scale_fill_manual(values  = color_palette) +
    scale_shape_manual(values = c(21, 22, 23, 24, 25, 7)) +
    scale_x_continuous(breaks = c(0, 2, 4, 6, 10, 24)) +
    scale_y_continuous(
      labels = label_math(10^.x),
      limits = c(0, 10)
    ) +
    labs(
      tag   = panel_label,
      x     = if (show_x_title) "Time (hours)" else NULL,
      y     = "CFU/mL (log₁₀)"
    ) +
    theme_tk +
    theme(
      plot.tag       = element_text(size = 9, face = "bold", hjust = 0),
      plot.tag.position = "topleft",
      legend.position   = if (show_legend) "top" else "none",
      legend.box        = "horizontal",
      axis.title.x      = if (show_x_title) element_text() else element_blank(),
      axis.text.x       = if (show_x_title) element_text() else element_blank(),
      axis.ticks.x      = if (show_x_title) element_line() else element_blank()
    ) +
    guides(
      color = "none",
      fill  = "none",
      shape = guide_legend(
        override.aes = list(fill  = rev(shape_colors),
                            color = rev(shape_colors)),
        title = "ATM-AVI (mg/L)",
        nrow  = 1
      )
    )
  
  return(p)
}

# ============================================================
# DRAW BOTH PANELS
# ============================================================
dummy_row <- summary_df %>%
  filter(name == "BC-8mg/L") %>%        # Copy structure from existing 8mg/L rows
  mutate(
    name       = "PHBC-8mg/L",          # New dummy condition name
    phage      = "Phage",               # Assign to phage group
    antibiotic = factor(8),
    mean_log   = NA,                    # All NA — nothing will be plotted
    sd_log     = NA,
    se_log     = NA,
    log_reduction = NA
  )

summary_df_balanced <- bind_rows(summary_df, dummy_row)

p_ab    <- plot_timekill(summary_df_balanced,
                         phage_group   = "No Phage",
                         panel_label   = "",
                         show_legend   = T,
                         show_x_title  = T)

p_phage <- plot_timekill(summary_df_balanced,
                         phage_group   = "Phage",
                         panel_label   = "",
                         show_legend   = T,   # Legend on bottom panel
                         show_x_title  = T)

# ============================================================
# COMBINE WITH PATCHWORK
# ============================================================

p_combined <- p_ab / p_phage +
  plot_layout(heights = c(1, 1)) &
  theme(plot.margin = margin(5, 10, 5, 10))

p_combined

ggsave("./figures/time_kill.pdf",width=10,height=7,device="pdf")
```

---
```{r}
# ============================================================
# TIME TO THRESHOLD METRICS
# ============================================================

thresholds <- c(1, 2, 3)   # log10 reductions to detect

time_to_kill <- long %>%
  filter(!is.na(log_cfu)) %>%
  group_by(name, phage, antibiotic, sample_id) %>%
  arrange(time) %>%
  mutate(
    log_reduction_rep = log10(initial_inoculum) - log_cfu
  ) %>%
  
  # For each threshold, find first timepoint where it is crossed
  summarise(
    time_1log = min(time[log_reduction_rep >= 1], default = Inf),
    time_2log = min(time[log_reduction_rep >= 2], default = Inf),
    time_3log = min(time[log_reduction_rep >= 3], default = Inf),
    max_kill  = max(log_reduction_rep, na.rm = TRUE),
    time_max  = time[which.max(log_reduction_rep)],
    .groups = "drop"
  ) %>%
  group_by(name, phage, antibiotic) %>%
  summarise(
    mean_time_1log = mean(time_1log[is.finite(time_1log)], na.rm = TRUE),
    mean_time_2log = mean(time_2log[is.finite(time_2log)], na.rm = TRUE),
    mean_time_3log = mean(time_3log[is.finite(time_3log)], na.rm = TRUE),
    pct_reach_1log = mean(is.finite(time_1log)) * 100,
    pct_reach_2log = mean(is.finite(time_2log)) * 100,
    pct_reach_3log = mean(is.finite(time_3log)) * 100,
    mean_max_kill  = mean(max_kill, na.rm = TRUE),
    mean_time_max  = mean(time_max, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    across(starts_with("mean_time"), 
           ~ifelse(is.nan(.) | is.infinite(.), NA, .)),
    across(starts_with("mean_time"),
           ~ifelse(is.na(.), ">24h", paste0(., "h")))
  )

time_to_kill %>%
  select(name, mean_time_1log, mean_time_2log, mean_time_3log,
         pct_reach_3log, mean_max_kill, mean_time_max) %>%
  arrange(mean_time_max) %>%
  kable(
    digits = 1,
    col.names = c("Condition",
                  "Time to 1-log kill",
                  "Time to 2-log kill",
                  "Time to 3-log kill",
                  "% reaching 3-log",
                  "Max log reduction",
                  "Time of max kill (h)"),
    caption = "Pharmacodynamic kill metrics per condition"
  )

```

```{r}

bubble_df <- long %>%
  filter(antibiotic %in% c(0,0.5, 1, 2, 4)) %>%
  filter(!is.na(log_cfu), time > 0) %>%
  mutate(log_reduction = log10(initial_inoculum) - log_cfu) %>%
  group_by(name, phage, time, antibiotic) %>%
  summarise(
    mean_reduction = mean(log_reduction, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    group_type = ifelse(phage == "Phage", "Phage + AB", "AB Only")
  )

plot_bubble <- function(data,
                        group,           # "AB Only" or "Phage + AB"
                        panel_label,     # "(A) ..." or "(B) ..."
                        show_legend  = FALSE,
                        show_x_title = FALSE) {

  df <- data %>% filter(group_type == group)

  ggplot(df, aes(x = time, y = factor(antibiotic))) +
    geom_point(aes(size = abs(mean_reduction),
                   fill = mean_reduction),
               shape = 21, color = "white", stroke = 0.5) +
    geom_text(aes(label = round(mean_reduction, 1)),
              size = 2.5, color = "black", fontface = "bold") +
    scale_fill_gradient2(
      low      = "#00FF9C",
      mid      = "#F39C12",
      high     = "#F93827",
      midpoint = 0,
      limits   = range(data$mean_reduction, na.rm = TRUE),  # Shared scale
      name     = "log10 reduction"
    ) +
    scale_size_continuous(range = c(3, 14), guide = "none") +
    scale_x_continuous(breaks = c(2, 4, 6, 10, 24)) +
    labs(
      tag = panel_label,
      x   = if (show_x_title) "Time (hours)" else NULL,
      y   = "ATM-AVI (mg/L)"
    ) +
    theme_tk +
    theme(
      plot.tag          = element_text(size = 9, face = "bold"),
      plot.tag.position = "topleft",
      legend.position   = if (show_legend) "top" else "none",
      legend.box = "horizontal",
      axis.title.x      = if (show_x_title) element_text() else element_blank(),
      axis.text.x       = if (show_x_title) element_text() else element_blank(),
      axis.ticks.x      = if (show_x_title) element_line()  else element_blank()
    )
}

p_bubble_ab    <- plot_bubble(bubble_df,
                              group        = "AB Only",
                              panel_label  = "",
                              show_legend  = T,
                              show_x_title = FALSE)

p_bubble_phage <- plot_bubble(bubble_df,
                              group        = "Phage + AB",
                              panel_label  = "",
                              show_legend  = F,
                              show_x_title = TRUE)

p_bubble_combined <- p_bubble_ab / p_bubble_phage +
  plot_layout(heights = c(1, 1), guides = "collect") &
  theme(plot.margin = margin(5, 10, 5, 10),  legend.position = "top")

p_bubble_combined

```

```{r}
# Δlog10 per timepoint × concentration
synergy_bubble_df <- long %>%
  filter(antibiotic %in% c(0, 0.5, 1, 2, 4),
         !is.na(log_cfu), time > 0) %>%
  mutate(
    log_reduction = log10(initial_inoculum) - log_cfu,
    group_type    = ifelse(phage == "Phage", "Phage + AB", "AB Only")
  ) %>%
  group_by(time, antibiotic, group_type) %>%
  summarise(mean_red = mean(log_reduction, na.rm = TRUE),
            .groups  = "drop") %>%
  pivot_wider(names_from  = group_type,
              values_from = mean_red) %>%
  mutate(
    delta      = `Phage + AB` - `AB Only`,   # Synergy only
    sig_border = case_when(                   # From your Dunn results
      time %in% c(4, 6) ~ "Peak",
      TRUE               ~ "Other"
    )
  )

p_synergy_bubble <- ggplot(
  synergy_bubble_df,
  aes(x = factor(time), y = factor(antibiotic))
) +
  # Significance border on peak timepoints
  annotate("rect",
           xmin = 1.5, xmax = 3.4,
           ymin = 0.5, ymax = 5.4,
           fill = "#8E44AD", alpha = 0.06,
           color = "#8E44AD", linetype = "dashed") +
  annotate("text", x = 2.5, y = 5.5,
           label = "Peak synergy (4h ≈ 6h, Dunn p < 0.001)",
           size = 3, color = "black", fontface = "bold") +

  geom_point(aes(size = abs(delta), fill = delta),
             shape = 21, color = "white", stroke = 0.6) +
  geom_text(aes(label = round(delta, 1),
                color = abs(delta) > 1.5),
            size = 2.8, fontface = "bold") +


  scale_fill_gradient2(
    low      = "#3498DB",    # Blue = phage worse (antagonism)
    mid      = "#ECF0F1",    # White = no synergy
    high     = "#E74C3C",    # Red = strong synergy
    midpoint = 0,
    name     = "Δlog₁₀\n(synergy)",
    limits   = c(-3, 6)
  ) +
  scale_color_manual(values = c("TRUE"  = "white",
                                 "FALSE" = "black"),
                     guide  = "none") +
  scale_size_continuous(range = c(3, 14), guide = "none") +
  scale_x_discrete(labels = paste0(c(2, 4, 6, 10, 24), "h")) +
  scale_y_discrete(labels = c("0" = "Phage\nonly",
                               "0.5" = "0.5",
                               "1" = "1",
                               "2" = "2",
                               "4" = "4")) +
  labs(
    x = "",
    y = "ATM-AVI (mg/L)",
  ) +
  theme_tk +
  theme(plot.tag = element_text(face = "bold", size = 12),
        plot.margins = c(5,5,5,5),
        legend.position=c(0.96,0.6),
        legend.background = element_blank())

p_synergy_bubble 
```



---
```{r stats_setup}
library(rstatix)
library(ggpubr)
library(knitr)

timepoint_df <- long %>%
  filter(antibiotic %in% c(0.5, 1, 2, 4),
         !is.na(log_cfu),
         time > 0) %>%
  mutate(
    group_type    = ifelse(phage == "Phage", "Phage + AB", "AB Only"),
    log_reduction = log10(initial_inoculum) - log_cfu
  )
```

### Test 1: Wilcoxon Test at Each Timepoint per Concentration

> Tests whether phage+AB reduces CFU significantly vs AB alone at each hour.

```{r stats_wilcoxon}
wilcox_by_time <- timepoint_df %>%
  group_by(time) %>%
  group_modify(~ {
    x <- .x %>% filter(group_type == "Phage + AB") %>% pull(log_reduction)
    y <- .x %>% filter(group_type == "AB Only")    %>% pull(log_reduction)

    result <- tryCatch(
      wilcox.test(x, y, exact = FALSE, correct = FALSE),
      warning = function(w) wilcox.test(x, y, exact = FALSE, correct = FALSE),
      error   = function(e) list(statistic = NA, p.value = NA)
    )

    tibble(
      n_phage_ab            = length(x),
      n_ab_only             = length(y),
      mean_reduction_phageAB = round(mean(x, na.rm = TRUE), 2),
      mean_reduction_ABonly  = round(mean(y, na.rm = TRUE), 2),
      delta_log              = round(mean(x, na.rm = TRUE) -
                                     mean(y, na.rm = TRUE), 2),
      W                      = as.numeric(result$statistic),
      p                      = result$p.value
    )
  }) %>%
  ungroup() %>%
  adjust_pvalue(method = "BH") %>%
  add_significance(p.col = "p.adj") %>%
  mutate(
    time_label     = paste0(time, "h"),
    interpretation = case_when(
      delta_log >= 2  ~ "Strong synergy",
      delta_log >= 1  ~ "Moderate synergy",
      delta_log >= 0  ~ "Mild synergy",
      TRUE            ~ "Antagonism"
    )
  ) %>%
  select(time_label, n_phage_ab, n_ab_only,
         mean_reduction_phageAB, mean_reduction_ABonly,
         delta_log, W, p, p.adj, p.adj.signif, interpretation)

wilcox_by_time %>%
  kable(
    digits = 3,
    col.names = c("Time", "n (Phage+AB)", "n (AB Only)",
                  "Mean log-red (Phage+AB)", "Mean log-red (AB Only)",
                  "Δlog₁₀", "W", "p-value", "p-adj (BH)", "Sig.",
                  "Interpretation"),
    caption = "Phage+AB vs AB alone at each timepoint (pooled across concentrations)"
  )

write_tsv(wilcox_by_time,"./export/wilcox_by_time_pooled_concentrations.tsv")
```

---

### Test 2: Effect Size (Cohen's d)

> Quantifies HOW MUCH phage reduces CFU, not just whether the difference is significant.

```{r stats_effectsize}
# ============================================================
# TEST 2: Kruskal-Wallis — does synergy differ across timepoints?
# ============================================================

# Calculate delta per matched pair (same antibiotic + replicate number)
delta_by_time <- timepoint_df %>%
  # Extract replicate number from sample_id
  mutate(rep = str_extract(sample_id, "_[0-9]+$") %>% str_remove("_")) %>%
  group_by(time, antibiotic, rep, group_type) %>%
  summarise(mean_red = mean(log_reduction, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = group_type, values_from = mean_red) %>%
  mutate(delta = `Phage + AB` - `AB Only`) %>%
  filter(!is.na(delta))

delta_by_time <- delta_by_time %>%
  mutate(time_f = factor(time))

# Check before running Kruskal
cat("Rows in delta_by_time:", nrow(delta_by_time), "\n")
cat("Timepoints present:", unique(delta_by_time$time), "\n")
cat("Groups per timepoint:\n")
print(table(delta_by_time$time))

# Only run if more than one timepoint has data
if (n_distinct(delta_by_time$time) > 1 &
    nrow(delta_by_time) > n_distinct(delta_by_time$time)) {

  kw_time <- kruskal_test(delta_by_time, delta ~ time_f)
  cat("Kruskal-Wallis: H =", round(kw_time$statistic, 3),
      "| df =", kw_time$df,
      "| p =",  round(kw_time$p, 4), "\n")

  dunn_time <- dunn_test(delta_by_time, delta ~ time_f,
                         p.adjust.method = "BH") %>%
    add_significance() %>%
    select(group1, group2, statistic, p, p.adj, p.adj.signif) %>%
    arrange(p.adj)

  dunn_time %>%
    kable(
      digits    = 3,
      col.names = c("Time 1", "Time 2", "Z", "p", "p-adj", "Sig."),
      caption   = "Dunn post-hoc: which timepoints differ in synergy magnitude?"
    )

} else {
  cat("⚠ Not enough matched pairs for Kruskal-Wallis.\n")
  cat("  Check that both Phage+AB and AB Only exist at each timepoint.\n")
  print(delta_by_time)
}

write_tsv(dunn_time,"./export/dunn_post_hoc_synergy_mag.tsv")
```

```{r}
# ============================================================
# DATA FOR PLOT
# ============================================================

# Delta summary per timepoint
delta_summary <- delta_by_time %>%
  group_by(time) %>%
  summarise(
    mean_delta = mean(delta, na.rm = TRUE),
    sd_delta   = sd(delta,   na.rm = TRUE),
    se_delta   = sd_delta / sqrt(n()),
    n          = n(),
    .groups    = "drop"
  )

# Compact letter display from Dunn results
# Groups: 2h=a, 4h&6h=b (not different), 10h&24h=a
cld_labels <- tibble(
  time       = c(2,    4,    6,    10,   24),
  cld        = c("a",  "b",  "b",  "a",  "a"),   # same letter = not significantly different
  y_pos      = delta_summary$mean_delta +
               delta_summary$se_delta + 0.4
)

# Phase annotations
phases <- tibble(
  xmin  = c(0.5, 1.5, 3.5),
  xmax  = c(1.5, 3.5, 5.5),
  label = c("Early\n(building)", "Peak synergy\n(4h ≈ 6h)", "Declining\n(regrowth)"),
  col   = c("#F39C12",            "#8E44AD",                  "#E74C3C")
)

# ============================================================
# PLOT
# ============================================================

p_timepoint_synergy <- ggplot(delta_summary, aes(x = factor(time), y = mean_delta)) +

  # Phase background shading
  annotate("rect",
           xmin  = phases$xmin, xmax = phases$xmax,
           ymin  = -Inf,        ymax = Inf,
           fill  = phases$col,  alpha = 0.1) +
  annotate("text",
           x     = rowMeans(cbind(phases$xmin, phases$xmax)),
           y     = Inf,
           label = phases$label,
           vjust = 1.3, size = 4,
           color = phases$col, fontface = "bold") +

  # Reference lines
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  geom_hline(yintercept = c(1, 2), linetype = "dotted",
             color = "#8E44AD", alpha = 0.5, linewidth = 0.6) +
  annotate("text", x = 0.55, y = c(1.08, 2.08),
           label = c("≥1 log synergy", "≥2 log synergy"),
           size = 2.8, color = "#8E44AD", hjust = 0) +

  # Individual observations (jitter)
  geom_jitter(data = delta_by_time,
              aes(x = factor(time), y = delta),
              width = 0.1, size = 3,
              alpha = 0.5, color = "black",fill="#605B51", shape = 21) +

  # Connecting line through means
  geom_line(aes(group = 1),
            color = "#8E44AD", linewidth = 0.8,
            linetype = "solid", alpha = 0.6) +

  # Error bars
  geom_errorbar(aes(ymin = mean_delta - se_delta,
                    ymax = mean_delta + se_delta),
                width = 0.15, linewidth = 0.8,
                color = "#8E44AD") +

  # Mean points — highlight 4h & 6h
  geom_point(aes(fill = factor(time) %in% c("4", "6")),
             shape = 21, size = 5,
             color = "white", stroke = 1) +
  scale_fill_manual(values = c("TRUE"  = "#8E44AD",
                                "FALSE" = "#AAB7B8"),
                    guide = "none") +

  # Compact letter display (CLD)
  geom_text(data = cld_labels,
            aes(x = factor(time), y = y_pos, label = cld),
            size = 4.5, fontface = "bold", color = "black",
            inherit.aes = FALSE) +


  scale_x_discrete(labels = paste0(c(2, 4, 6, 10, 24), "h")) +
  scale_y_continuous(breaks = seq(-4, ceiling(max(delta_by_time$delta,
                                                   na.rm = TRUE)) + 1, 1)) +
  labs(
    x        = "Time (hours)",
    y        = "Δlog₁₀ CFU/mL (synergy)"
  ) + theme_tk 
```

```{r}
library(patchwork)

# Row 1: Kill curves side by side
row1 <- (p_ab | p_phage) + plot_layout(guides = "collect") & theme(legend.position = "top")

# Row 2: Synergy bubble + Timepoint stats
row2 <- (p_synergy_bubble / p_timepoint_synergy) +  plot_layout(heights = c(1, 1))

# Combine
final_figure <- row1 / row2 +
  plot_layout(heights = c(1, 1.5)) +
  plot_annotation(tag_levels =  "A") &
  theme(plot.margin = margin(1, 1, 1, 1),
        axis.title.y.left = element_text(size = 10),plot.tag = element_text(size=11,face="bold"))

ggsave("./figures/time_kill_combined2.png",width=10,height=13,device="png")

```




