---
title: "IMC Analysis: Phage-Antibiotic Synergy"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 150)
```

```{r libraries}
library(tidyverse)
library(patchwork)
library(ggrepel)
library(rstatix)
library(knitr)
library(scales)

# ============================================================
# PARAMETERS
# ============================================================

imc_file <- "../../time_kill_data/heat.tsv"

# Shared theme (consistent with time-kill figures)
theme_tk <- theme_bw(base_size = 12) +
  theme(
    panel.grid.minor  = element_blank(),
    strip.background  = element_rect(fill = "#F0F0F0"),
    legend.position   = "right",
    plot.title        = element_text(face = "bold", size = 13)
  )

# Colors per condition group
imc_colors <- c(
  "Bacteria"         = "#2C3E50",
  "Phage alone"      = "#2980B9",
  "AB 0.5 mg/L"      = "#FAD7A0",
  "AB 1 mg/L"        = "#F39C12",
  "AB 2 mg/L"        = "#E67E22",
  "AB 8 mg/L"        = "#E74C3C",
  "AB 16 mg/L"       = "#922B21",
  "Phage + 0.5 mg/L" = "#D7BDE2",
  "Phage + 1 mg/L"   = "#9B59B6",
  "Phage + 2 mg/L"   = "#8E44AD",
  "Media control"    = "#BDC3C7"
)
```

## Data Import & Cleaning

```{r data_prep}
raw_imc <- read.delim(imc_file, stringsAsFactors = FALSE, check.names = FALSE)

# Rename columns for easier handling
colnames(raw_imc) <- c("well", "annotation", "roi_start", "roi_end",
                        "total_heat_J", "metabolic_rate_uW",
                        "max_metabolic_rate_uW", "TTA_h", "TTP_h")

# Parse annotation to clean condition labels
imc <- raw_imc %>%
  mutate(
    # Clean annotation (remove duplicate text from column)
    annotation = str_trim(str_split_i(annotation, " CTL", 1)),
    
    # Map to readable condition
    condition = case_when(
      annotation == "CTL BACT"    ~ "Bacteria",
      annotation == "CTL EC10P1"  ~ "Phage alone",
      annotation == "CTL MH"      ~ "Media control",
      annotation == "CTL BACT 0.5"~ "AB 0.5 mg/L",
      annotation == "CTL BACT 1"  ~ "AB 1 mg/L",
      annotation == "CTL BACT 2"  ~ "AB 2 mg/L",
      annotation == "CTL BACT 8"  ~ "AB 8 mg/L",
      annotation == "CTL BACT 16" ~ "AB 16 mg/L",
      annotation == "10P1 0.5MG"  ~ "Phage + 0.5 mg/L",
      annotation == "10P1 1MG"    ~ "Phage + 1 mg/L",
      annotation == "10P1 2MG"    ~ "Phage + 2 mg/L",
      TRUE                        ~ annotation
    ),
    
    # Group: AB only, Phage+AB, Controls
    group = case_when(
      str_detect(condition, "Phage \\+") ~ "Phage + AB",
      str_detect(condition, "AB")        ~ "AB only",
      condition == "Phage alone"         ~ "Phage alone",
      TRUE                               ~ "Control"
    ),
    
    # Extract AB concentration for matched comparisons
    ab_conc = case_when(
      str_detect(condition, "0\\.5") ~ 0.5,
      str_detect(condition, "1 mg")  ~ 1,
      str_detect(condition, "2 mg")  ~ 2,
      str_detect(condition, "8 mg")  ~ 8,
      str_detect(condition, "16 mg") ~ 16,
      TRUE                           ~ 0
    ),
    
    # Handle Inf TTA (bacteria never activated)
    TTA_h_clean   = ifelse(is.infinite(TTA_h), NA, TTA_h),
    suppressed     = is.infinite(TTA_h),
    
    # Factor ordering
    condition = factor(condition, levels = names(imc_colors))
  )

# Summary statistics per condition
imc_summary <- imc %>%
  group_by(condition, group, ab_conc) %>%
  summarise(
    n                  = n(),
    mean_TTP           = mean(TTP_h,              na.rm = TRUE),
    sd_TTP             = sd(TTP_h,                na.rm = TRUE),
    se_TTP             = sd_TTP / sqrt(n),
    mean_TTA           = mean(TTA_h_clean,         na.rm = TRUE),
    sd_TTA             = sd(TTA_h_clean,           na.rm = TRUE),
    mean_total_heat    = mean(total_heat_J,         na.rm = TRUE),
    sd_total_heat      = sd(total_heat_J,           na.rm = TRUE),
    se_total_heat      = sd_total_heat / sqrt(n),
    mean_max_rate      = mean(max_metabolic_rate_uW,na.rm = TRUE),
    sd_max_rate        = sd(max_metabolic_rate_uW,  na.rm = TRUE),
    se_max_rate        = sd_max_rate / sqrt(n),
    pct_suppressed     = mean(suppressed) * 100,
    .groups = "drop"
  )

cat("Conditions:\n")
print(imc_summary %>% select(condition, group, n, pct_suppressed))
```

---

## Plot A: Time to Peak (TTP) — Primary Metric

> Later peak = more suppression of bacterial metabolic activity.
> Phage+AB should show significantly delayed TTP vs AB alone or phage alone.

```{r plot_TTP, fig.height=5, fig.width=10}
# Reference lines
bact_TTP <- imc_summary %>%
  filter(condition == "Bacteria") %>%
  pull(mean_TTP)

phage_TTP <- imc_summary %>%
  filter(condition == "Phage alone") %>%
  pull(mean_TTP)

p_TTP <- imc_summary %>%
  filter(condition != "Media control") %>%
  ggplot(aes(y = reorder(as.character(condition), mean_TTP),
             x = mean_TTP, color = condition)) +

  # Reference lines: bacteria and phage alone
  geom_vline(xintercept = bact_TTP,
             linetype = "dashed", color = "#2C3E50", alpha = 0.6) +
  geom_vline(xintercept = phage_TTP,
             linetype = "dashed", color = "#2980B9", alpha = 0.6) +
  annotate("text", x = bact_TTP + 0.2, y = 0.6,
           label = "Bacteria\nalone", size = 2.8,
           color = "#2C3E50", hjust = 0) +
  annotate("text", x = phage_TTP + 0.2, y = 0.6,
           label = "Phage\nalone", size = 2.8,
           color = "#2980B9", hjust = 0) +

  # Individual points
  geom_jitter(data = imc %>% filter(condition != "Media control"),
              aes(y = condition, x = TTP_h, color = condition),
              width = 0, height = 0.15,
              size = 2.5, alpha = 0.5, shape = 16,
              inherit.aes = FALSE) +

  # Mean ± SE
  geom_errorbar(aes(xmin = mean_TTP - se_TTP,
                    xmax = mean_TTP + se_TTP),
                width = 0.25, linewidth = 0.8) +
  geom_point(aes(fill = condition),
             shape = 21, size = 4,
             color = "white", stroke = 0.8) +

  # Group labels on right
  geom_text(aes(x = 19, label = group),
            size = 3, hjust = 1, color = "grey40") +

  scale_color_manual(values = imc_colors, guide = "none") +
  scale_fill_manual(values  = imc_colors, guide = "none") +
  scale_y_discrete() +   
  scale_x_continuous(breaks = seq(0, 18, 2), limits = c(0, 20)) +
  labs(
    title    = "Time to Peak Metabolic Activity (TTP)",
    subtitle = "Later peak = greater suppression of bacterial growth",
    x        = "Time to Peak (hours)",
    y        = NULL,
    tag      = "A"
  ) +
  theme_tk +
  theme(plot.tag = element_text(face = "bold", size = 12))

p_TTP
```

---

## Plot B: Total Heat Production

> Lower total heat = less overall metabolic activity = more bacterial suppression.

```{r plot_heat, fig.height=5, fig.width=10}
bact_heat <- imc_summary %>%
  filter(condition == "Bacteria") %>%
  pull(mean_total_heat)

p_heat <- imc_summary %>%
  filter(condition != "Media control") %>%
 ggplot(aes(y = reorder(as.character(condition), -mean_total_heat),
             x = mean_total_heat, color = condition)) +

  geom_vline(xintercept = bact_heat, linetype = "dashed",
             color = "#2C3E50", alpha = 0.6) +
  annotate("text", x = bact_heat + 0.02, y = 0.6,
           label = "Bacteria\nalone", size = 2.8,
           color = "#2C3E50", hjust = 0) +

  # Individual points
  geom_jitter(data = imc %>% filter(condition != "Media control"),
              aes(y = condition, x = total_heat_J, color = condition),
              width = 0, height = 0.15,
              size = 2.5, alpha = 0.5, shape = 16,
              inherit.aes = FALSE) +

  geom_errorbar(aes(xmin = mean_total_heat - se_total_heat,
                    xmax = mean_total_heat + se_total_heat),
                width = 0.25, linewidth = 0.8) +
  geom_point(aes(fill = condition),
             shape = 21, size = 4,
             color = "white", stroke = 0.8) +

  scale_color_manual(values = imc_colors, guide = "none") +
  scale_fill_manual(values  = imc_colors, guide = "none") +
  scale_y_discrete()+
  labs(
    title    = "Total Heat Production (Q)",
    subtitle = "Lower Q = less overall bacterial metabolic activity",
    x        = "Total Heat (J)",
    y        = NULL,
    tag      = "B"
  ) +
  theme_tk +
  theme(plot.tag = element_text(face = "bold", size = 12))

p_heat
```

---

## Plot C: Matched Comparison — Phage+AB vs AB alone

> Direct synergy test at matched concentrations (0.5, 1, 2 mg/L).
> Both TTP and Total Heat shown side by side.

```{r plot_matched, fig.height=6, fig.width=11}
# Only matched concentrations (phage+AB exists for 0.5, 1, 2 mg/L)
matched_df <- imc %>%
  filter(ab_conc %in% c(0.5, 1, 2),
         group %in% c("AB only", "Phage + AB")) %>%
  mutate(ab_label = paste0(ab_conc, " mg/L"))

matched_summary <- imc_summary %>%
  filter(ab_conc %in% c(0.5, 1, 2),
         group %in% c("AB only", "Phage + AB")) %>%
  mutate(ab_label = paste0(ab_conc, " mg/L"))

# TTP matched
p_match_TTP <- ggplot(matched_summary,
                      aes(x = ab_label, y = mean_TTP,
                          color = group, group = group)) +
  geom_jitter(data = matched_df,
              aes(x = ab_label, y = TTP_h, color = group),
              position = position_jitterdodge(jitter.width = 0.1,
                                              dodge.width  = 0.5),
              size = 2.5, alpha = 0.5) +
  geom_line(position = position_dodge(0.5),
            linewidth = 0.8, linetype = "dashed", alpha = 0.6) +
  geom_errorbar(aes(ymin = mean_TTP - se_TTP,
                    ymax = mean_TTP + se_TTP),
                position = position_dodge(0.5),
                width = 0.15, linewidth = 0.8) +
  geom_point(aes(fill = group),
             position = position_dodge(0.5),
             shape = 21, size = 4, color = "white", stroke = 0.8) +
  scale_color_manual(values = c("AB only"    = "#E67E22",
                                "Phage + AB" = "#8E44AD")) +
  scale_fill_manual(values  = c("AB only"    = "#E67E22",
                                "Phage + AB" = "#8E44AD")) +
  labs(x = "ATM-AVI Concentration",
       y = "Time to Peak (h)",
       color = NULL, fill = NULL,
       tag = "C") +
  theme_tk +
  theme(plot.tag       = element_text(face = "bold", size = 12),
        legend.position = "bottom")

# Total heat matched
p_match_heat <- ggplot(matched_summary,
                       aes(x = ab_label, y = mean_total_heat,
                           color = group, group = group)) +
  geom_jitter(data = matched_df,
              aes(x = ab_label, y = total_heat_J, color = group),
              position = position_jitterdodge(jitter.width = 0.1,
                                              dodge.width  = 0.5),
              size = 2.5, alpha = 0.5) +
  geom_line(position = position_dodge(0.5),
            linewidth = 0.8, linetype = "dashed", alpha = 0.6) +
  geom_errorbar(aes(ymin = mean_total_heat - se_total_heat,
                    ymax = mean_total_heat + se_total_heat),
                position = position_dodge(0.5),
                width = 0.15, linewidth = 0.8) +
  geom_point(aes(fill = group),
             position = position_dodge(0.5),
             shape = 21, size = 4, color = "white", stroke = 0.8) +
  scale_color_manual(values = c("AB only"    = "#E67E22",
                                "Phage + AB" = "#8E44AD")) +
  scale_fill_manual(values  = c("AB only"    = "#E67E22",
                                "Phage + AB" = "#8E44AD")) +
  labs(x = "ATM-AVI Concentration",
       y = "Total Heat (J)",
       color = NULL, fill = NULL,
       tag = "D") +
  theme_tk +
  theme(plot.tag        = element_text(face = "bold", size = 12),
        legend.position = "bottom")

p_match_TTP | p_match_heat
```

---

## Plot E: Multi-metric IMC Summary Heatmap

> Overview of all 4 metrics across all conditions — useful for supplementary.

```{r plot_heatmap, fig.height=6, fig.width=10}
# Scale each metric 0-1 for comparable heatmap
heatmap_df <- imc_summary %>%
  filter(condition != "Media control") %>%
  select(condition, group,
         `TTP (h)`        = mean_TTP,
         `Total Heat (J)` = mean_total_heat,
         `Max Rate (µW)`  = mean_max_rate) %>%
  pivot_longer(
    cols      = c(`TTP (h)`, `Total Heat (J)`, `Max Rate (µW)`),  # match exact renamed cols
    names_to  = "metric",
    values_to = "value"
  ) %>%
  group_by(metric) %>%
  mutate(
    scaled = (value - min(value, na.rm = TRUE)) /
             (max(value, na.rm = TRUE) - min(value, na.rm = TRUE)),
    metric = factor(metric, levels = c("TTP (h)", "Total Heat (J)", "Max Rate (µW)"))
  ) %>%
  ungroup()



ggplot(heatmap_df,
       aes(x = metric,
           y = fct_reorder(condition, value),
           fill = scaled)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = round(value, 1),
                color = scaled > 0.6),
            size = 3, fontface = "bold") +
  facet_grid(group ~ ., scales = "free_y", space = "free") +
  scale_fill_gradient2(
    low      = "#2ECC71",
    mid      = "#F39C12",
    high     = "#C0392B",
    midpoint = 0.5,
    name     = "Scaled\nvalue",
    #labels   = c("Low", "", "High")
  ) +
  scale_color_manual(values = c("TRUE"  = "white",
                                 "FALSE" = "black"),
                     guide  = "none") +
  labs(
    title    = "IMC Summary — All Metrics",
    subtitle = "Values scaled 0–1 per metric for comparison | Red = high activity, Green = suppressed",
    x        = NULL, y = NULL,
    tag      = "E"
  ) +
  theme_tk +
  theme(
    plot.tag         = element_text(face = "bold", size = 12),
    axis.text.x      = element_text(angle = 30, hjust = 1),
    strip.text.y     = element_text(angle = 0),
    panel.spacing    = unit(0.3, "cm")
  )
```

---

## Statistical Tests

```{r stats_imc}
# Wilcoxon: Phage+AB vs AB alone at matched concentrations, per metric
imc_stats <- imc %>%
  filter(ab_conc %in% c(0.5, 1, 2),
         group %in% c("AB only", "Phage + AB")) %>%
  mutate(ab_label = paste0(ab_conc, " mg/L")) %>%
  pivot_longer(cols = c(TTP_h, total_heat_J, max_metabolic_rate_uW),
               names_to  = "metric",
               values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(metric, ab_label) %>%
  group_modify(~ {
    x <- .x %>% filter(group == "Phage + AB") %>% pull(value)
    y <- .x %>% filter(group == "AB only")    %>% pull(value)
    if (length(x) < 2 | length(y) < 2) {
      return(tibble(W = NA, p = NA))
    }
    result <- tryCatch(
      wilcox.test(x, y, exact = FALSE, correct = FALSE),
      error = function(e) list(statistic = NA, p.value = NA)
    )
    tibble(
      mean_phage_ab = round(mean(x, na.rm = TRUE), 3),
      mean_ab_only  = round(mean(y, na.rm = TRUE), 3),
      W             = as.numeric(result$statistic),
      p             = result$p.value
    )
  }) %>%
  ungroup() %>%
  adjust_pvalue(method = "BH") %>%
  add_significance(p.col = "p.adj") %>%
  mutate(metric = recode(metric,
    "TTP_h"                  = "Time to Peak (h)",
    "total_heat_J"           = "Total Heat (J)",
    "max_metabolic_rate_uW"  = "Max Rate (µW)"
  ))

imc_stats %>%
  arrange(metric, ab_label) %>%
  kable(
    digits    = 4,
    col.names = c("Metric", "AB Conc", "Mean (Phage+AB)",
                  "Mean (AB only)", "W", "p", "p-adj", "Sig."),
    caption   = "Wilcoxon: Phage+AB vs AB alone per metric and concentration"
  )
```

---

## Summary Table

```{r summary_table}
imc_summary %>%
  filter(condition != "Media control") %>%
  mutate(
    `TTP mean ± SE`        = paste0(round(mean_TTP, 1),        " ± ", round(se_TTP, 1)),
    `Total Heat mean ± SE` = paste0(round(mean_total_heat, 3), " ± ", round(se_total_heat, 3)),
    `Max Rate mean ± SE`   = paste0(round(mean_max_rate, 1),   " ± ", round(se_max_rate, 1)),
    `% Suppressed`         = paste0(round(pct_suppressed), "%")
  ) %>%
  select(condition, group, n,
         `TTP mean ± SE`, `Total Heat mean ± SE`,
         `Max Rate mean ± SE`, `% Suppressed`) %>%
  arrange(group, condition) %>%
  kable(caption = "IMC summary statistics per condition (mean ± SE)")
```

---

## Key Findings

```{r findings, echo=FALSE, results='asis'}
bact_TTP_val  <- imc_summary %>% filter(condition == "Bacteria")   %>% pull(mean_TTP)  %>% round(1)
phage_TTP_val <- imc_summary %>% filter(condition == "Phage alone")%> pull(mean_TTP)  %>% round(1)
combo_TTP_val <- imc_summary %>% filter(condition == "Phage + 2 mg/L") %>% pull(mean_TTP) %>% round(1)

cat(glue::glue("
**Time to Peak:**
- Bacteria alone: **{bact_TTP_val}h** (earliest, fastest growth)
- Phage alone: **{phage_TTP_val}h** (delayed — phage suppresses then regrows)
- Phage + 2 mg/L: **{combo_TTP_val}h** (most delayed — sustained suppression)

**Interpretation:**
- Sub-MIC antibiotic alone does not delay TTP vs bacteria alone
- Phage alone delays TTP by ~{round(phage_TTP_val - bact_TTP_val, 1)}h
- Phage + AB delays TTP further — consistent with synergistic suppression
- AB 8-16 mg/L fully suppresses activity (Inf TTA) — above MIC, expected
"))
```
