---
title: "Interaction-Defined Transcriptional Responses at 4 h"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi = 300
)
```

```{r libraries}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(scales)
```

---

# Load data

```{r load}
df     <- read.csv("./export/interaction/combo_therapy_responses.csv",    stringsAsFactors = FALSE)
int4   <- read.csv("./export/__response_calss_Interaction-defined_(Non-additive)_4h.csv",
                   stringsAsFactors = FALSE)

df4    <- df %>% filter(time_point == "4h")
```

---

# Shared theme and palettes

```{r theme}
# ── Condition colours ────────────────────────────────────────────────────────
cond_pal <- c(
  "Phage"            = "#4CAF50",
  "Antibiotic"       = "#FF9800",
  "Combination"      = "#7B1FA2",
  "Expected additive"= "#90A4AE",   # dashed reference
  "Interaction"      = "#E53935"
)

# ── Cluster palette (7 clusters) ────────────────────────────────────────────
cluster_pal <- c(
  "Anaerobic/metabolic escape"  = "#D32F2F",
  "Proteostasis"                = "#1565C0",
  "Oxidative defence"           = "#E65100",
  "Outer membrane & permeability" = "#2E7D32",
  "Motility & chemotaxis"       = "#6A1B9A",
  "Surface & biofilm"           = "#4E342E",
  "Metabolic & other"           = "#78909C"
)

# ── Shared ggplot theme ──────────────────────────────────────────────────────
theme_pub <- function(base = 10) {
  theme_classic(base_size = base) %+replace%
    theme(
      plot.title       = element_text(face = "bold", size = base + 1, hjust = 0),
      plot.subtitle    = element_text(colour = "grey45", size = base - 1, hjust = 0),
      axis.text        = element_text(colour = "black"),
      panel.grid.major = element_line(colour = "grey93", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      legend.title     = element_text(face = "bold", size = base - 1),
      legend.text      = element_text(size  = base - 1),
      strip.background = element_rect(fill = "grey95", colour = NA),
      strip.text       = element_text(face = "bold", size = base - 1)
    )
}
```

---

# Panel A — Interaction model schematic using real gene examples

Two illustrative genes demonstrate what a *negative* and a *positive* interaction
coefficient mean using the actual data.

```{r panel-A-data}
# torA: strong negative interaction (antibiotic-induced program blunted by phage)
# fliA: positive interaction (antibiotic-repressed gene partially rescued by phage)
eg_genes <- c("torA", "fliA")

pA_df <- df4 %>%
  filter(gene.y %in% eg_genes) %>%
  select(gene = gene.y,
         Phage        = lfc_phage,
         Antibiotic   = lfc_ab,
         Combination  = lfc_combo,
         Interaction  = lfc_interaction) %>%
  distinct(gene, .keep_all = TRUE) %>%
  mutate(
    Expected_additive = Phage + Antibiotic,
    gene_label = case_when(
      gene == "torA" ~ "italic(torA)~(negative~interaction)",
      gene == "fliA" ~ "italic(fliA)~(positive~interaction)"
    )
  )

# Long format for the four observed conditions
pA_long <- pA_df %>%
  pivot_longer(cols = c(Phage, Antibiotic, Combination),
               names_to  = "condition",
               values_to = "lfc") %>%
  mutate(condition = factor(condition,
                            levels = c("Phage","Antibiotic","Combination")))
```

```{r panel-A-plot, fig.width = 9, fig.height = 4.5}
pA <- ggplot(pA_long, aes(x = condition, y = lfc, group = gene)) +

  # Zero reference
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.4) +

  # Expected additive: open diamond
  geom_point(
    data = pA_df %>%
      pivot_longer(Expected_additive, names_to = "condition", values_to = "lfc") %>%
      mutate(condition = "Combination"),
    aes(x = condition, y = lfc),
    shape = 23, size = 4, fill = "white",
    colour = cond_pal["Expected additive"], stroke = 1.2,
    inherit.aes = FALSE
  ) +

  # Dashed line from Antibiotic to Expected additive
  geom_segment(
    data = pA_df,
    aes(x = "Antibiotic", xend = "Combination",
        y = Antibiotic,   yend = Expected_additive),
    linetype = "dashed", colour = cond_pal["Expected additive"],
    linewidth = 0.7, inherit.aes = FALSE
  ) +

  # Observed lines
  geom_line(colour = "grey50", linewidth = 0.6) +

  # Observed points
  geom_point(aes(colour = condition), size = 3.5) +

  # Arrow showing interaction gap (Combination → Expected)
  geom_segment(
    data = pA_df,
    aes(x    = 2.85, xend = 2.85,
        y    = Combination,
        yend = Expected_additive),
    arrow = arrow(ends = "both", length = unit(0.07, "in"),
                  type = "closed"),
    colour    = cond_pal["Interaction"],
    linewidth = 0.8,
    inherit.aes = FALSE
  ) +

  # Interaction label
  geom_text(
    data = pA_df,
    aes(x = 3.05,
        y = (Combination + Expected_additive) / 2,
        label = paste0("Δ = ", round(Interaction, 2))),
    colour   = cond_pal["Interaction"],
    size     = 3.2,
    fontface = "bold",
    hjust    = 0,
    inherit.aes = FALSE
  ) +

  facet_wrap(~ gene_label, scales = "free_y",
             labeller = label_parsed) +

  scale_colour_manual(
    values = cond_pal[c("Phage","Antibiotic","Combination")],
    name   = "Condition"
  ) +
  scale_x_discrete(expand = expansion(add = c(0.5, 0.9))) +

  labs(
    title    = "A  Interaction model — illustrative gene examples",
    subtitle = "Open diamond = expected additive (phage + antibiotic LFC); filled point = observed combination;\ncoloured arrow = interaction coefficient (Δ)",
    x        = NULL,
    y        = expression(log[2]~"fold change")
  ) +
  theme_pub() +
  theme(legend.position = "right",
        panel.grid.major.x = element_blank())

pA
```

---

# Panel B — Full landscape of interaction-defined genes (4 h)

All 113 interaction-defined genes sorted by interaction coefficient, coloured
by functional cluster.

```{r panel-B-data}
# Interaction rows from the annotated file
inter_base <- int4 %>%
  filter(group2 == "I") %>%
  select(gene_name, functional_group, lfc_interaction = LFC) %>%
  distinct(gene_name, .keep_all = TRUE)

inter_base <- int4 %>%
  filter(group2 == "I") %>%
  select(gene_name, functional_group, lfc_interaction = LFC) %>%
  distinct(gene_name, .keep_all = TRUE)

# Manual cluster assignment (based on gene function review)
manual_clusters <- tribble(
  ~gene_name,  ~cluster,
  # Anaerobic/metabolic escape
  "torA","Anaerobic/metabolic escape", "torC","Anaerobic/metabolic escape",
  "torD","Anaerobic/metabolic escape", "hycC","Anaerobic/metabolic escape",
  "hycE","Anaerobic/metabolic escape", "hycF","Anaerobic/metabolic escape",
  "hycG","Anaerobic/metabolic escape", "hycH","Anaerobic/metabolic escape",
  "fdhF","Anaerobic/metabolic escape", "dmsD","Anaerobic/metabolic escape",
  "cadA","Anaerobic/metabolic escape", "lysU","Anaerobic/metabolic escape",
  "asnB","Anaerobic/metabolic escape", "yecD","Anaerobic/metabolic escape",
  "yhjX","Anaerobic/metabolic escape", "nepI","Anaerobic/metabolic escape",
  "creA","Anaerobic/metabolic escape",
  # Proteostasis
  "dnaK","Proteostasis", "dnaJ","Proteostasis",
  "groL","Proteostasis", "groS","Proteostasis", "htpG","Proteostasis",
  "hslU","Proteostasis", "hslV","Proteostasis",
  "dnaA","Proteostasis", "recF","Proteostasis",
  "ybeZ","Proteostasis", "ybeY","Proteostasis",
  "pspA","Proteostasis", "rho","Proteostasis",
  # Oxidative defence
  "ahpC","Oxidative defence", "ahpF","Oxidative defence",
  "cnoX","Oxidative defence", "dsbC","Oxidative defence",
  "yqhD","Oxidative defence", "ydjA","Oxidative defence",
  "aldA","Oxidative defence", "pdxJ","Oxidative defence",
  # Outer membrane & permeability
  "ompF","Outer membrane & permeability",
  "chiP","Outer membrane & permeability",
  "znuA","Outer membrane & permeability", "znuC","Outer membrane & permeability",
  "pstS","Outer membrane & permeability", "feoA","Outer membrane & permeability",
  "emrD","Outer membrane & permeability", "fadL","Outer membrane & permeability",
  # Motility & chemotaxis
  "fliA","Motility & chemotaxis", "flgB","Motility & chemotaxis",
  "flgD","Motility & chemotaxis", "flgG","Motility & chemotaxis",
  "tap","Motility & chemotaxis",  "cheR","Motility & chemotaxis",
  "trg","Motility & chemotaxis",
  # Surface & biofilm
  "fimB","Surface & biofilm", "fimI","Surface & biofilm",
  "mlrA","Surface & biofilm", "wzc","Surface & biofilm",
  "csgC","Surface & biofilm", "hns","Surface & biofilm",
  "ppdA","Surface & biofilm", "mltF","Surface & biofilm",
  "rdcB","Surface & biofilm", "sfmD","Surface & biofilm"
)

pB_df <- inter_base %>%
  left_join(manual_clusters, by = "gene_name") %>%
  mutate(cluster = replace_na(cluster, "Metabolic & other"),
         cluster = factor(cluster, levels = names(cluster_pal))) %>%
  arrange(lfc_interaction) %>%
  mutate(rank = row_number())

# Genes to label
label_set <- c(
  "torA","torC","torD","hycF","hycE",
  "dnaK","groL","hslU",
  "ahpF","ahpC",
  "ompF","cadA","znuA",
  "fliA","tap","flgB",
  "pspA","recF"
)
```

```{r panel-B-plot, fig.width = 10, fig.height = 6}
pB <- ggplot(pB_df, aes(x = rank, y = lfc_interaction, colour = cluster)) +

  # Zero reference
  geom_hline(yintercept = 0, colour = "grey50",
             linewidth = 0.5, linetype = "dashed") +

  # Shaded regions
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "#FFEBEE", alpha = 0.25) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = 0,    ymax = Inf,
           fill = "#E8F5E9", alpha = 0.25) +

  # Stems
  geom_segment(aes(x = rank, xend = rank, y = 0, yend = lfc_interaction),
               linewidth = 0.45, alpha = 0.7) +

  # Points
  geom_point(size = 2.2) +

  # Labels
  geom_label_repel(
    data        = filter(pB_df, gene_name %in% label_set),
    aes(label   = gene_name),
    size        = 2.8,
    fontface    = "italic",
    fill        = "white",
    label.size  = 0.2,
    label.padding = unit(0.15, "lines"),
    box.padding = 0.3,
    max.overlaps = 30,
    segment.size   = 0.3,
    segment.colour = "grey55",
    show.legend  = FALSE
  ) +

  # Region annotations
  annotate("text", x = 5, y = -2.6,
           label = "Antibiotic-induced programs\nblunted by phage",
           colour = "#B71C1C", size = 2.9, fontface = "italic", hjust = 0) +
  annotate("text", x = 75, y = 2.6,
           label = "Antibiotic-repressed programs\nrescued by phage",
           colour = "#1B5E20", size = 2.9, fontface = "italic", hjust = 0) +

  scale_colour_manual(values = cluster_pal,
                      name   = "Functional cluster",
                      guide  = guide_legend(
                        override.aes = list(size = 3),
                        keywidth = 0.6, keyheight = 0.6
                      )) +
  scale_y_continuous(breaks = seq(-3, 3, by = 1),
                     limits = c(-3.1, 3.1)) +
  scale_x_continuous(expand = expansion(mult = 0.02)) +

  labs(
    title    = "B  Interaction-defined genes at 4 h (n = 113)",
    subtitle = "Ranked by interaction coefficient; colour = functional cluster",
    x        = "Genes ranked by interaction coefficient",
    y        = expression("Interaction coefficient (\u0394log"[2]*"FC)")
  ) +
  theme_pub() +
  theme(
    axis.text.x    = element_blank(),
    axis.ticks.x   = element_blank(),
    legend.position = "right"
  )

pB
```

---

# Panel C — Resistance architecture context

A data-driven dot plot placing the outer membrane, efflux, and mar/acr
genes in their resistance-layer context, showing that none carry a
significant interaction term — contrasting with the interaction-defined
clusters.

```{r panel-C-data}
# Genes that define the resistance architecture context
arch_genes <- c(
  # Outer membrane layer
  "ompF","ompF_2","ompC","ompA","ompW","ompR","ompX","ompN","ompG",
  # Efflux
  "acrA","acrB","acrR","acrZ","acrD","acrE","acrF","tolC",
  # mar regulon
  "marA","marR","marB","marC"
)

arch_layer <- tribble(
  ~gene_y,  ~layer,
  "ompF",   "Outer membrane\n(OmpF frameshifted)",
  "ompF_2", "Outer membrane\n(OmpF frameshifted)",
  "ompC",   "Outer membrane\n(OmpF frameshifted)",
  "ompA",   "Outer membrane\n(OmpF frameshifted)",
  "ompW",   "Outer membrane\n(OmpF frameshifted)",
  "ompR",   "Outer membrane\n(OmpF frameshifted)",
  "ompX",   "Outer membrane\n(OmpF frameshifted)",
  "ompN",   "Outer membrane\n(OmpF frameshifted)",
  "ompG",   "Outer membrane\n(OmpF frameshifted)",
  "acrA",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrB",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrR",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrZ",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrD",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrE",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "acrF",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "tolC",   "Efflux\n(AcrAB-TolC / AcrEF)",
  "marA",   "mar regulon",
  "marR",   "mar regulon",
  "marB",   "mar regulon",
  "marC",   "mar regulon"
)

pC_df <- df4 %>%
  filter(gene.y %in% arch_genes) %>%
  select(gene_y = gene.y,
         Phage        = lfc_phage,
         Antibiotic   = lfc_ab,
         Combination  = lfc_combo,
         Interaction  = lfc_interaction,
         sig_interaction, response_class,
         padj_phage, padj_ab, padj_interaction) %>%
  distinct(gene_y, .keep_all = TRUE) %>%
  left_join(arch_layer, by = "gene_y") %>%
  pivot_longer(cols = c(Phage, Antibiotic, Combination, Interaction),
               names_to  = "condition",
               values_to = "lfc") %>%
  mutate(
    condition   = factor(condition,
                         levels = c("Phage","Antibiotic","Combination","Interaction")),
    sig_inter   = sig_interaction,
    gene_y      = factor(gene_y,
                         levels = rev(arch_genes[arch_genes %in% unique(pC_df_tmp <- df4 %>%
                                                                           filter(gene.y %in% arch_genes) %>%
                                                                           pull(gene.y))])),
    layer       = factor(layer, levels = unique(arch_layer$layer))
  )

# Reorder gene_y by layer then by phage lfc
gene_order <- df4 %>%
  filter(gene.y %in% arch_genes) %>%
  distinct(gene.y, .keep_all = TRUE) %>%
  left_join(arch_layer, by = c("gene.y" = "gene_y")) %>%
  arrange(layer, lfc_phage) %>%
  pull(gene.y) %>%
  rev()

pC_df <- pC_df %>%
  mutate(gene_y = factor(gene_y, levels = gene_order))
```

```{r panel-C-plot, fig.width = 10, fig.height = 7}

# Significance annotation data (for interaction column only)
sig_annot <- df4 %>%
  filter(gene.y %in% arch_genes) %>%
  distinct(gene.y, .keep_all = TRUE) %>%
  mutate(
    sig_label = case_when(
      padj_interaction < 0.001 ~ "***",
      padj_interaction < 0.01  ~ "**",
      padj_interaction < 0.05  ~ "*",
      TRUE                     ~ ""
    ),
    condition = factor("Interaction",
                       levels = c("Phage","Antibiotic","Combination","Interaction")),
    gene_y = factor(gene.y, levels = gene_order)
  ) %>%
  left_join(arch_layer, by = c("gene.y" = "gene_y"))

pC <- ggplot(pC_df,
             aes(x = condition, y = gene_y, fill = lfc)) +
  geom_tile(colour = "white", linewidth = 0.5) +

  # Significance stars in interaction column
  geom_text(
    data    = sig_annot,
    aes(x   = condition, y = gene_y, label = sig_label),
    colour  = "black", size = 3.5, fontface = "bold",
    inherit.aes = FALSE
  ) +

  facet_grid(layer ~ ., scales = "free_y",
             space = "free_y", switch = "y") +

  # Vertical divider before Interaction column
  geom_vline(xintercept = 3.5, colour = "grey30",
             linewidth = 0.8) +

  scale_fill_gradientn(
    colours = c("#1565C0","#64B5F6","#E3F2FD",
                "white",
                "#FFCDD2","#E57373","#B71C1C"),
    values  = rescale(c(-6, -2, -0.5, 0, 0.5, 2, 6)),
    limits  = c(-6, 6),
    oob     = squish,
    name    = expression(log[2]*"FC"),
    guide   = guide_colourbar(
      barwidth = 0.5, barheight = 7,
      title.position = "top", title.hjust = 0.5
    )
  ) +

  scale_x_discrete(
    labels = c("Phage","Antibiotic","Combination",
               expression(bold("Interaction \u0394")))
  ) +

  labs(
    title    = "C  Outer membrane, efflux, and mar regulon — no interaction (4 h)",
    subtitle = "Resistance-layer genes: phage-driven remodelling without combination-specific interaction\n* padj < 0.05 in interaction column (ompF only; frameshifted — no functional protein)",
    x        = NULL,
    y        = NULL
  ) +
  theme_pub() +
  theme(
    axis.text.y        = element_text(size = 8.5, face = "italic"),
    axis.text.x        = element_text(size = 9),
    strip.text.y.left  = element_text(angle = 0, hjust = 1,
                                      size = 8, face = "bold"),
    strip.placement    = "outside",
    panel.grid         = element_blank(),
    panel.spacing      = unit(0.2, "lines"),
    legend.position    = "right"
  )

pC
```

---

# Panel D — Slope graph: observed vs expected additive for focal genes

Shows the gap between expected additive and observed combination for the
most mechanistically important genes across four functional clusters.

```{r panel-D-data}
focal_genes <- tribble(
  ~gene,   ~cluster,
  # Anaerobic/metabolic escape
  "torA",  "Anaerobic/metabolic\nescape",
  "torC",  "Anaerobic/metabolic\nescape",
  "torD",  "Anaerobic/metabolic\nescape",
  "hycF",  "Anaerobic/metabolic\nescape",
  "hycE",  "Anaerobic/metabolic\nescape",
  # Proteostasis
  "dnaK",  "Proteostasis",
  "groL",  "Proteostasis",
  "hslU",  "Proteostasis",
  "hslV",  "Proteostasis",
  # Oxidative defence
  "ahpF",  "Oxidative defence",
  "ahpC",  "Oxidative defence",
  # Outer membrane / context
  "ompF",  "Outer membrane\n& acid stress",
  "cadA",  "Outer membrane\n& acid stress"
)

pD_df <- df4 %>%
  filter(gene.y %in% focal_genes$gene) %>%
  distinct(gene.y, .keep_all = TRUE) %>%
  select(gene = gene.y,
         lfc_phage, lfc_ab, lfc_combo, lfc_interaction) %>%
  mutate(lfc_expected = lfc_phage + lfc_ab) %>%
  left_join(focal_genes, by = "gene") %>%
  mutate(
    cluster = factor(cluster, levels = unique(focal_genes$cluster)),
    gene    = factor(gene,    levels = rev(unique(focal_genes$gene)))
  )

# Long for slope: three x positions — Phage, Antibiotic, Expected, Observed
pD_long <- pD_df %>%
  select(gene, cluster, lfc_phage, lfc_ab, lfc_expected, lfc_combo) %>%
  pivot_longer(
    cols      = c(lfc_phage, lfc_ab, lfc_expected, lfc_combo),
    names_to  = "condition",
    values_to = "lfc"
  ) %>%
  mutate(
    condition = factor(condition,
                       levels = c("lfc_phage","lfc_ab","lfc_expected","lfc_combo"),
                       labels = c("Phage","Antibiotic",
                                  "Expected\nadditive","Observed\ncombination")),
    line_type = ifelse(condition %in% c("Expected\nadditive","Observed\ncombination"),
                       "focal", "mono")
  )
```

```{r panel-D-plot}
pD <- ggplot(pD_long,
             aes(x = condition, y = lfc, group = gene,
                 colour = cluster)) +

  geom_hline(yintercept = 0, colour = "grey55",
             linewidth = 0.4, linetype = "dashed") +

  # Lines connecting all four points
  geom_line(aes(linetype = line_type), linewidth = 0.6, alpha = 0.75) +

  # Points
  geom_point(aes(shape = condition), size = 2.5) +

  # Expected vs Observed gap shading
  geom_segment(
    data = pD_df,
    aes(x    = 3.5, xend = 3.5,
        y    = lfc_expected,
        yend = lfc_combo,
        colour = cluster),
    linewidth = 2.5, alpha = 0.2,
    inherit.aes = FALSE
  ) +

  # Gene labels at observed combination point
  geom_text_repel(
    data        = filter(pD_long, condition == "Observed\ncombination"),
    aes(label   = gene),
    size        = 2.8,
    fontface    = "italic",
    nudge_x     = 0.15,
    direction   = "y",
    hjust       = 0,
    segment.size   = 0.3,
    segment.colour = "grey60",
    show.legend = FALSE
  ) +

  facet_wrap(~ cluster, nrow = 1, scales = "free_y") +

  scale_colour_manual(values = cluster_pal, guide = "none") +
  scale_linetype_manual(values = c("focal" = "solid", "mono" = "solid"),
                        guide  = "none") +
  scale_shape_manual(
    values = c("Phage" = 16, "Antibiotic" = 16,
               "Expected\nadditive" = 5, "Observed\ncombination" = 16),
    guide  = "none"
  ) +
  scale_x_discrete(expand = expansion(add = c(0.4, 1.0))) +

  labs(
    title    = "D  Expected vs observed combination response — focal gene clusters (4 h)",
    subtitle = "Diamond = expected additive (phage LFC + antibiotic LFC); filled point = observed combination;\nshaded bar = interaction gap",
    x        = NULL,
    y        = expression(log[2]~"fold change")
  ) +
  theme_pub() +
  theme(
    panel.grid.major.x = element_blank(),
    strip.text         = element_text(size = 8.5)
  )

pD
```

---

# Compose final figure

```{r compose, fig.width = 14, fig.height = 20}
# fig_final <-
#   pA /
#   pB /
#   (pC | pD) +

fig_final <-
  pB / pD / pC +
  plot_annotation(
    tag_levels = "A",
    title      = "Interaction-defined transcriptional responses reveal phage-mediated\ndismantling of stress tolerance in a multi-drug-resistant isolate",
    caption    = paste0(
      "LFC: log\u2082 fold change vs untreated. Interaction coefficient (\u0394): ",
      "departure of observed combination from expected additive (phage + antibiotic LFC). ",
      "Isolate carries NDM-5, YRIK insertion in PBP3, and frameshifted ompF."
    ),
    theme = theme(
      plot.title   = element_text(face = "bold", size = 13, hjust = 0,
                                  lineheight = 1.2),
      plot.caption = element_text(colour = "grey50", size = 8, hjust = 0)
    )
  ) +
  plot_layout(heights = c(0.7, 0.9, 1.2))

fig_final
```

```{r save}
if (!dir.exists("./figures")) dir.create("./figures", recursive = TRUE)

ggsave("./figures/Figure_interaction_panels.pdf",
       fig_final, width = 14, height = 20,
       device = cairo_pdf)

ggsave("./figures/Figure_interaction_panels.png",
       fig_final, width = 14, height = 20,
       dpi = 300)

# Individual panels for journal submission flexibility
ggsave("./figures/PanelA_model.pdf",     pA, width = 9,  height = 4.5, device = cairo_pdf)
ggsave("./figures/PanelB_landscape.pdf", pB, width = 10, height = 6,   device = cairo_pdf)
ggsave("./figures/PanelC_resistance.pdf",pC, width = 10, height = 7,   device = cairo_pdf)
ggsave("./figures/PanelD_slopegraph.pdf",pD, width = 11, height = 7,   device = cairo_pdf)

message("All figures saved.")
```

---

```{r session}
sessionInfo()
```
